---
title: "PPOL 670 | Week 6 | Breakout 1"
subtitle: "Scraping Wikipedia Page on Presidential Debates" 
output: 
  html_document:
    highlight: breezedark
    theme: united
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r,include = F}
knitr::opts_chunk$set(warning=F,error=F,message = F)
```

<br><br><br>

# Overview

The following Wikipedia webpage contains information on past presidential debates. 

```{r,eval=F}
url <-  "https://en.wikipedia.org/wiki/United_States_presidential_debates"
```


Please use this url to answer the following questions. 

<br><br>

# Questions

<br>

#### (1) Download the wikipedia website using the provided `url`.

<br><br><br>

#### (2) Scrape all the text from the wiki page. Collapse content into a single string.

<br><br><br>

#### (3) Scrape all the references from the wikipedia page and store in a `tibble` data.frame.

<br><br><br>

#### (4)  Download and clean the table after the "Timeline" header on the page. Create a tidy data frame that only has the election date and the presidential candidates as variables. 

<br><br><br>
