---
title: "PPOL 670 | Week 4 | Breakout 2"
subtitle: "Joining and Shaping Data" 
output: 
  html_document:
    highlight: breezedark
    theme: united
---

```{r include = FALSE}
knitr::opts_chunk$set(warning=F,error=F,message = F)
# Load the tidyverse package
require(tidyverse)
```

## Instructions

The following code generates two datasets and assign the data to the respective object names `beertax` and `dogs`.


```{r data}
# Data Set 1
beertax <- 
  tibble::tribble(
    ~STATE, ~YEAR, ~beer_tax,
    "Montana",  2006,      0.27,
    "Maine",  2006,      0.28,
    "Arkansas",  2006,      0.09,
    "Kansas",  2006,      0.05,
    "New Mexico",  2006,      0.24,
    "Tennessee",  2006,      0.23,
    "Massachusetts",  2006,      0.28,
    "Connecticut",  2006,       0.3,
    "Minnesota",  2006,      0.28,
    "Florida",  2006,      0.15,
    "Oklahoma",  2006,      0.09,
    "Washington",  2006,      0.08,
    "Idaho",  2006,      0.16,
    "Colorado",  2006,      0.15,
    "North Carolina",  2006,       0.1,
    "Oregon",  2006,      0.29,
    "Virginia",  2006,      0.19,
    "California",  2006,      0.05,
    "South Carolina",  2006,      0.13,
    "West Virginia",  2006,      0.28,
    "Rhode Island",  2006,      0.15,
    "South Dakota",  2006,      0.27,
    "Alabama",  2006,      0.18,
    "New Hampshire",  2006,      0.19,
    "Maryland",  2006,      0.26,
    "Nevada",  2006,      0.16,
    "Indiana",  2006,       0.3,
    "Texas",  2006,       0.1,
    "New York",  2006,      0.15,
    "Arizona",  2006,      0.11,
    "Vermont",  2006,      0.19,
    "Utah",  2006,      0.23,
    "Wisconsin",  2006,      0.17,
    "Delaware",  2006,      0.29,
    "Missouri",  2006,      0.18,
    "Hawaii",  2006,      0.14,
    "Alaska",  2006,      0.08,
    "Illinois",  2006,      0.03,
    "Mississippi",  2006,      0.26,
    "Louisiana",  2006,      0.08
  )


# Dataset 2
dogs <- tibble::tribble(
            ~state, ~year, ~dogs_per_1000,
        "Colorado",  2006,          13.52,
      "New Mexico",  2006,          15.48,
      "New Jersey",  2006,           3.25,
   "West Virginia",  2006,           7.49,
  "South Carolina",  2006,          10.07,
   "New Hampshire",  2006,          12.82,
         "Alabama",  2006,           7.41,
         "Georgia",  2006,          13.47,
          "Nevada",  2006,           11.5,
        "Michigan",  2006,          11.24,
       "Tennessee",  2006,          10.02,
    "South Dakota",  2006,           9.85,
          "Hawaii",  2006,           12.9,
            "Ohio",  2006,           8.01,
    "Pennsylvania",  2006,          10.05,
      "Washington",  2006,          10.04,
           "Texas",  2006,          12.71,
         "Florida",  2006,          14.12,
         "Wyoming",  2006,           3.08,
        "Delaware",  2006,           1.18,
    "North Dakota",  2006,          12.17,
  "North Carolina",  2006,          10.09,
      "California",  2006,          10.65,
          "Alaska",  2006,          14.97,
        "New York",  2006,          13.13
  )
```

Please use the `_join()` functions from the tidyverse (`dplyr`) to complete each question. In addition, some of the questions require you to reshape the data. For these question, please use the `pivot_longer()` and `pivot_wider()` functions (`tidyr`). Both functions can be acquired by loading the `tidyverse` package



## Question 1


#### (a) Join `beertax` and `dogs` data so that only the states in `dogs` are retained? Assign to the object named `join1`.

```{r}

```

#### (b) How many entries for the `beer_tax` variable are missing from the `join1` data? Why are these values missing?

```{r}

```


#### (c) Join `beertax` and `dogs` data so that only the states in `beertax` are retained? Assign to the object named `join2`.

```{r}

```

#### (d) How many entries for the `dogs_per_1000` variable are missing from the `join2` data? Why are these values missing?

```{r}

```


# Question 2

#### (a) spread the `join2` data into a wide format. Where the `year` makes up the columns and `dogs_per_1000` populate the values for each of the spread out columns. Assign the output to the object `fulldat`.

```{r}

```


#### (b) using `replace_na()` replace all missing values with `-99` in `fulldat`. Overwrite the `fulldat` object to retain your changes. (Tip: use the documentation to learn how to use `?replace_na()`)

```{r}

```
