---
title: "PPOL 670 | Week 4 | Breakout 1"
subtitle: "Manipulating Conflict Event Data" 
output: 
  html_document:
    css: async-page-style.css
    highlight: breezedark
    theme: united
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r,include = F}
knitr::opts_chunk$set(error=F,warning = F)
```

# Overview

In this notebook, we'll hone our data manipulation skills by examining conflict event data generated by the [Armed Conflict Location & Event Data Project (ACLED)](https://acleddata.com/#/dashboard). The aim is to practice some of the tidyverse functions covered in the lecture. 

As always, we need to make sure we have the `tidyverse` packages installed and loaded. 

```{r}
# install.packages("tidyverse")  
require(tidyverse) 
```


# Data 

ACLED is a "disaggregated data collection, analysis, and crisis mapping project. ACLED collects the dates, actors, locations, fatalities, and modalities of all reported political violence and protest events across Africa, South Asia, Southeast Asia, the Middle East, Central Asia and the Caucasus, Latin America and the Caribbean, and Southeastern and Eastern Europe and the Balkans." For this exercise, we'll focus just on the data pertaining to _Africa_. For more information regarding these data, please consult the [ACLED methodology](https://acleddata.com/resources/methodology/).

```{r,warning=F,message=F}
acled <- read_csv("https://raw.githubusercontent.com/edunford/ppol670/gh-pages/Lectures/week_04/walkthrough/acled_africa.csv")
head(acled)
```


# Know your data

To understand a dataset, one needs to ask a lot of questions of it. To get a better feel for the ACLED data, let's explore the following questions:

#### 1. What's the temporal coverage of the data (i.e how many years does it span)?

```{r}

```


#### 2. How many events are recorded for each country? 

```{r}

```

#### 3. How many events are recorded for each year?

```{r}

```


#### 4. What's the most common event type in the data?

```{r}

```


#### 5. How many `sub_` event types are there for each event type?

```{r}

```


#### 6. Which countries had the highest number of reported fatalities? 

```{r}

```

