---
pagetitle: PPOL670 - Introduction to Data Science
output: 
  html_document:
    css: style.css
    theme: united
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
---

```{r,include=F}
# Packages for reading in the tables
require(tidyverse)
require(kableExtra)

# FUTURE UPDATES
# Add an additional data wrangling day (or extend the current line up) to include:
# - case_when()
# - janitor

# Add gghighlight() to data viz day. 
# Add interactive graphics

# No PDFs on assignments. 

# More in-class coding. Slow down so there can be more of this. 
```

<center><h1>Introduction to Data Science</h1></center>
<center><h2>PPOL670-01</h2></center>
<center>**Fall 2020**</center>
<center>**Georgetown University**</center>

<br>



# Course Outline {.tabset .tabset-fade .tabset-pills}

## Course Schedule

<br>
<br>


```{r,echo=F,message=F,results='asis'}
read_csv('Syllabus/fall-2020-timeline.csv') %>%
  # filter(Topic != "No Class") %>% 
  {.[is.na(.)]="";.} %>% 
  kable(., "html", booktabs = T,align = c('c','l','l')) %>%
  kable_styling(latex_options = c("striped", 
                                  "scale_down","repeat_header"),
                full_width = T,font_size = 20) %>% 
  column_spec(1,width = '.5in') %>% 
  column_spec(2,width = '2in') %>% 
  column_spec(3,width = '4in') 
```

---

<br>
<br>

## Virtual Classroom


<br> 


#### [Link to Virtual Zoom Classroom](https://georgetown.zoom.us/j/91598655229)
  
<br>
  
#### [Virtual Office Hours](https://georgetown.zoom.us/j/96444963432) (Tuesdays 9am - 11am)

  
<br>

Recurrent Zoom link can also be found on Canvas. Please contact the professor/TA through Slack if any of the Zoom links break. 

<br>
<br>
<br>
<hr>
<br>

## Syllabus

<br>
<br>

#### Access an [**online version**](Syllabus/PPOL670-Introduction-to-DS_Syllabus.html)/[**`.pdf` version**](Syllabus/PPOL670-Introduction-to-DS_Syllabus.pdf).

<br>
<br>

---

<br>
<br>

## Readings

<br>

#### List of [**all required and suggested readings**](Readings/reading-list.html)

<br>
<br>

---

<br>
<br>


## Project

<br>

Data science is an applied field and therefore, it is important that you understand how to conduct a complete analysis from collecting data, to cleaning and analyzing it, to presenting your findings. Toward this end, students are required to complete an independent data science project, _applying concepts learned throughout the course_. The project is composed of three parts: a 2 page project proposal, an in-class presentation, and a 12-page project report.

**_More information regarding the final project will be circulated during class on Week 8_**



<!-- <br> -->

<!-- **Overview** -->

<!-- <center>[**Project Overview**](Assignments/Project/project_overview.pdf)</center> -->

<!-- <br> -->

<!-- **Rubrics** -->

<!-- <center>[**Presentation Assessment Sheet**](Assignments/Project/project_presentation_assessment_sheet.pdf)</center> -->

<!-- <center>[**Final Project Rubric**](Assignments/Project/final_project_rubric.pdf)</center> -->

<!-- <br> -->

<!-- **Examples of Successful Projects** -->

<!-- - [Kristina Costa - "Analyzing FEMA Pre-Disaster Mitigation Grants"](Assignments/Project/Costa_Kristina_Example_Final_Report.pdf) -->
<!-- - [Liping Wang - "Female Labor Force Participation"](Assignments/Project/Wang_Liping_Example_Final_Report.pdf) -->



<br>
<br>

---

<br>
<br>


## Installation

<br>
<br>

The following are installation instructions for `R` and `RStudio`.

<br>
<br>

**`R` Software**

To install `R`, download `R` from CRAN via the following:
 
- Windows: https://cran.r-project.org/bin/windows/base/ (Links to an external site.)
- Mac: https://cran.r-project.org/bin/macosx/
 
To install `RStudio`, download from the following (scroll to the bottom):

- https://www.rstudio.com/products/rstudio/download/ (Links to an external site.)

Video walkthroughs:

- [Video Walkthrough](https://www.dropbox.com/s/7af77pyb1wh224w/R-startup.mp4?dl=0) - [_Put together by former TA (Nathan Lovin)_]
- RStudio Videos
  - [Install `R`](https://player.vimeo.com/video/203516510)
  - [Install `RStudio`](https://player.vimeo.com/video/203516968)

<!-- <br> -->
<!-- <br> -->

<!-- **Git**  -->

<!-- - Windows/Mac: https://git-scm.com/downloads (Links to an external site.) -->
<!--     + Note that git comes pre-installed on the macOS operating system.  -->

<!-- Video walkthroughs: -->

<!--  - [TA Video Walkthrough: Using Git with RStudio](https://www.dropbox.com/s/qdw38hbvy0kj630/git.mp4?dl=0) - [_Put together by former TA (Nathan Lovin)_] -->


<br>
<br>

---

<br>
<br>


## Contact

<br>

#### [Eric Dunford](http://ericdunford.com/) (Professor)

- **Office**: Bedroom (formerly 404 Old North)
- **Office Hours**: Tuesdays 9:00am to 11:00am (EST) or by appointment
- **Email**: [eric.dunford@georgetown.edu](mailto:eric.dunford@georgetown.edu)

<br>

#### [Maddie Pickens](http://maddieclairepickens.com/) (Teaching Assistant)

- **Office Hours**: [Book via Calendarly](https://calendly.com/maddiepickens)
- **Email**: [mp1595@georgetown.edu](mailto:mp1595@georgetown.edu)

<br>

#### Class

- **Slack**: class communications will primarily take place through [Slack](https://get.slack.help/hc/en-us/articles/218080037-Getting-started-for-new-members). Please follow the [**_invite link_**](https://join.slack.com/t/ppol670-fall-2020/shared_invite/zt-gnbnemyd-9GSNAaEfXtSUgzsnH3KHKA) to be added to the class Slack channel.

<br>
<br>

---

<br>
<br>


# Lecture Materials
_Asynchronous lecture materials will go live approximately one week prior to the scheduled synchronous meeting date._

<br>
<br>

### _Week 1_: **Work Flow and Reproducibility**

<br>

- [**Lecture Slides**](Lectures/week_01/lecture-week-01.html)

  - **_Table of Contents_**
    + [About this Course](http://ericdunford.com/ppol670/Lectures/week_01/lecture-week-01.html#3)
    + [Reproducibility](http://ericdunford.com/ppol670/Lectures/week_01/lecture-week-01.html#14)
    + [Interacting with `R`](http://ericdunford.com/ppol670/Lectures/week_01/lecture-week-01.html#14)

<br>


### _Week 2_: **Introduction to Programming in `R`**


<br>

- [**Asynchronous Material**](Lectures/week_02/asynchronous_lecture_material/week-02-async-material.html)

- [**Lecture Slides**](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html)

  - **_Table of Contents_**
    - [Objects](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#3)
    - [Data Structures](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#20)
    - [Indexing](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#28)
    - [Operators](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#57)
    - [Functions](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#70)
    - [Packages](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#76)
    - [Importing and Exporting Data](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#85)
    - [But where is my data exactly?](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#100)

- **Practice**
  + In breakout rooms, let's [practice creating an R project, subsetting and creating data structures in R, and importing/exporting data.](Lectures/week_02/Practice/week-02_practice.html)

<br>

### _Week 3_: **Reproducibility in Practice**

<br>

- [**Asynchronous Material**](Lectures/week_03/asynchronous_lecture_material/week-03-async-material.html)


- [**Lecture Slides**](Lectures/week_03/lecture-week03-reproducibility-in-practice.html)

  - **Table of Contents**
    + [Iteration](http://ericdunford.com/ppol670/Lectures/week_03/lecture-week03-reproducibility-in-practice.html#3)
    + [Control Statements](http://ericdunford.com/ppol670/Lectures/week_03/lecture-week03-reproducibility-in-practice.html#22)
    + [`R` Markdown](http://ericdunford.com/ppol670/Lectures/week_03/lecture-week03-reproducibility-in-practice.html#26)
    + [Generating Reproducible Examples](http://ericdunford.com/ppol670/Lectures/week_03/lecture-week03-reproducibility-in-practice.html#50)

- **Practice**:
  + _Breakout 1_: [Piecing together the code](Lectures/week_03/Practice/breakout01_piecing-together-code.html)
  + <a href="Lectures/week_03/Practice/rmarkdown_walkthrough.Rmd.zip" download> R Markdown Walkthrough (download `.Rmd`) </a>
  + _Breakout 2_: [Generate a Reproducible Example](Lectures/week_03/Practice/breakout02_reprex_practice.html)
  

<br>

### _Week 4_: **Data Wrangling in R**


<br>

- [**Asynchronous Material**](Lectures/week_04/asynchronous_lecture_material/week4-async-material.html)


- [**Lecture Slides**](Lectures/week_04/lecture-week-04_data-wrangling.html)
  - **Table of Contents**
    + [Data Wrangling](http://ericdunford.com/ppol670/Lectures/week_04/lecture-week-04_data-wrangling.html#3)
    + [Tidy Data](http://ericdunford.com/ppol670/Lectures/week_04/lecture-week-04_data-wrangling.html#9)
    + [Manipulating Data](http://ericdunford.com/ppol670/Lectures/week_04/lecture-week-04_data-wrangling.html#17)
    + [Piping](http://ericdunford.com/ppol670/Lectures/week_04/lecture-week-04_data-wrangling.html#60)
    + [Joining Data](http://ericdunford.com/ppol670/Lectures/week_04/lecture-week-04_data-wrangling.html#67)
    + [Reshaping Data](http://ericdunford.com/ppol670/Lectures/week_04/lecture-week-04_data-wrangling.html#85)
    
- **Practice**:
  + _Breakout 1_: <a href="Lectures/week_04/walkthrough/breakout01_manipulating-conflict-data.Rmd.zip" download> <strong>Manipulating conflict event data </strong> </a>
  + _Breakout 2_: <a href="Lectures/week_04/walkthrough/breakout02_joining-reshaping.Rmd.zip" download> <strong> Practice joining and reshaping data</strong> </a>

<br>

### _Week 5_: **Data Visualization**

<br>

- [**Asynchronous Material**](Lectures/week_05/asynchronous_lecture_material/week5-async-material.html)


<!-- - **Table of Contents** -->
<!--   + [The Components of Data Visualization](http://ericdunford.com/ppol670/Lectures/week_05/lecture-week-05_data-visualization.html#4) -->
<!--   + [Grammar of Graphics](http://ericdunford.com/ppol670/Lectures/week_05/lecture-week-05_data-visualization.html#17) -->
<!--   + [Dates](http://ericdunford.com/ppol670/Lectures/week_05/lecture-week-05_data-visualization.html#66) -->

<!-- - **Practice**: -->
<!--   - <a href="Lectures/week_05/R-code/dem-gdp-data.csv" download>Here</a> is a link to the data used in the visualizations in the lecture slides. -->
<!--   - Practice materials for visualizing data can be downloaded <a href="Lectures/week_05/Practice-Week-05-Visualization/practice_week-05.Rmd.zip" download>here</a>. Work through the questions then compare your results to the <a href="Lectures/week_05/Practice-Week-05-Visualization/practice_week-05_ANSWERS.Rmd.zip" download> answer key</a> -->

<br>

### _Week 6_: **Web Scrapping and APIs**

<!-- ### _Week 6_: [Web Scrapping and APIs](Lectures/week_06/lecture-week-06_webscraping-and-apis.html) -->

<!-- - **Table of Contents** -->
<!--   + [Loops and Functions (Refresher)](http://ericdunford.com/ppol670/Lectures/week_06/lecture-week-06_webscraping-and-apis.html#3) -->
<!--   + [Web Content](http://ericdunford.com/ppol670/Lectures/week_06/lecture-week-06_webscraping-and-apis.html#10) -->
<!--   + [Scraping Web Content](http://ericdunford.com/ppol670/Lectures/week_06/lecture-week-06_webscraping-and-apis.html#19) -->
<!--   + [Building a Webscraper](http://ericdunford.com/ppol670/Lectures/week_06/lecture-week-06_webscraping-and-apis.html#33) -->
<!--   + [APIs](http://ericdunford.com/ppol670/Lectures/week_06/lecture-week-06_webscraping-and-apis.html#45) -->

<!-- - **Practice**: -->
<!--   - The following [walkthrough](Lectures/week_06/week_06_practice/week-06_webscrapping-walkthrough.html) covers building a dataset of all bilaterial investment treaties from scratch using the techniques covered in lecture (<a href="Lectures/week_06/week_06_practice/week-06_webscrapping-walkthrough.Rmd.zip" download> Download .Rmd</a>). -->

<br>

### _Week 7_: **Geospatial Data**

<!-- - **Working with Spatial Data in `R`**  -->
<!--   - The following are materials provided by Trey Billing. He'll walk us through how to utilize spatial data in `R` and some common concepts we need to know when dealing with spatial data.  -->
<!--     + [Spatial Walkthrough](Lectures/week_13/spatial_walkthrough/walkthrough.html) -->
<!--     + [Walkthrough Materials](Lectures/week_13/spatial_walkthrough.zip) (`.zip`) -->

<br>

### _Week 8_: **Text as Data**

<!-- ### _Week 7_: [Text as Data](Lectures/week_07/lecture-week-07-text-as-data-ppol670.html) -->

<!-- - **Table of Contents** -->
<!--   + [Strings](http://ericdunford.com/ppol670/Lectures/week_07/lecture-week-07-text-as-data-ppol670.html#3) -->
<!--   + [Tidy Text](http://ericdunford.com/ppol670/Lectures/week_07/lecture-week-07-text-as-data-ppol670.html#20) -->
<!--   + [Sentiment Analysis](http://ericdunford.com/ppol670/Lectures/week_07/lecture-week-07-text-as-data-ppol670.html#49) -->
<!--   + [Topic Models](http://ericdunford.com/ppol670/Lectures/week_07/lecture-week-07-text-as-data-ppol670.html#64) -->


<!-- - **Practice**: -->
<!--   - The following provides a [walkthrough](Lectures/week_07/week_07_practice/walkthrough_text-as-data_ppol670.html) of the concepts covered in this lecture. The materials can be found  [here](Lectures/week_07/week_07_practice.zip) as a `.zip`. -->

<br>

### _Week 9_: **Introduction to Statistical Learning**

<!-- ### _Week 8_: [Introduction to Statistical Learning](Lectures/week_08/week-08-lecture-supervised-learning.html) -->
<!-- - **Table of Contents** -->
<!--   + [Statistical Learning](http://ericdunford.com/ppol670/Lectures/week_08/week-08-lecture-supervised-learning.html#3) -->
<!--   + [Supervised Learning](http://ericdunford.com/ppol670/Lectures/week_08/week-08-lecture-supervised-learning.html#13) -->
<!--   + [Cross-Validation](http://ericdunford.com/ppol670/Lectures/week_08/week-08-lecture-supervised-learning.html#32) -->
<!--   + [Caret](http://ericdunford.com/ppol670/Lectures/week_08/week-08-lecture-supervised-learning.html#56) -->
<!--   + [Pre-Processing Data](http://ericdunford.com/ppol670/Lectures/week_08/week-08-lecture-supervised-learning.html#64) -->

<br>

### _Week 10_: **Applications in Supervised Learning (Regression)**

<!-- ### _Week 9_: [Applications in Supervised Learning (Regression)](Lectures/week_09/lecture-week-09-applications-supervised-ml-regression-ppol670.html) -->
<!-- - **Table of Contents** -->
<!--   + [Regression Methods](http://ericdunford.com/ppol670/Lectures/week_09/lecture-week-09-applications-supervised-ml-regression-ppol670.html#3) -->
<!--   + [Regression Trees](http://ericdunford.com/ppol670/Lectures/week_09/lecture-week-09-applications-supervised-ml-regression-ppol670.html#33) -->
<!--   + [Bagging and Random Forest](http://ericdunford.com/ppol670/Lectures/week_09/lecture-week-09-applications-supervised-ml-regression-ppol670.html#48) -->

<!-- - **Practice**: -->
<!--   - The following provides a [walkthrough](Lectures/week_09/walkthrough/supervised-ml-regression-walkthrough.html) of the concepts covered in this lecture. The materials can be found  [here](Lectures/week_09/week_09-practice.zip) as a `.zip`. -->

<br>

### _Week 11_: **Applications in Supervised Learning (Classification)**

<!-- ### _Week 10_: [Applications in Supervised Learning (Classification)](Lectures/week_10/lecture-week-10-applications-supervised-ml-classification-ppol670.html) -->

<!-- - **Table of Contents**: -->
<!--   + [Classification](http://ericdunford.com/ppol670/Lectures/week_10/lecture-week-10-applications-supervised-ml-classification-ppol670.html#3) -->
<!--   + [Logistic Regression](http://ericdunford.com/ppol670/Lectures/week_10/lecture-week-10-applications-supervised-ml-classification-ppol670.html#16) -->
<!--   + [K-Nearest Neighbors](http://ericdunford.com/ppol670/Lectures/week_10/lecture-week-10-applications-supervised-ml-classification-ppol670.html#24) -->
<!--   + [Classification Trees](http://ericdunford.com/ppol670/Lectures/week_10/lecture-week-10-applications-supervised-ml-classification-ppol670.html#37) -->
<!--   + [Support Vector Machines](http://ericdunford.com/ppol670/Lectures/week_10/lecture-week-10-applications-supervised-ml-classification-ppol670.html#42) -->


<!-- - **Practice**: -->
<!--   - The following provides a [walkthrough](Lectures/week_10/walkthrough/supervised-ml-classification-walkthrough.html) of the concepts covered in this lecture. The materials can be found  [here](Lectures/week_10/week-10-walkthrough.zip) as a `.zip`. -->

<br>

### _Week 12_: **Interpretable Machine Learning**

<!-- ### _Week 12_: [Exploratory Data Analysis](Lectures/week_12/lecture-week-12-exploratory-data-analysis-ml-ppol670.html) -->

<!-- - **Table of Contents**: -->
<!--   + [Interpretable Machine Learning](http://ericdunford.com/ppol670/Lectures/week_12/lecture-week-12-exploratory-data-analysis-ml-ppol670.html#3) -->
<!--   + [Data Exploration in the Social Sciences](http://ericdunford.com/ppol670/Lectures/week_12/lecture-week-12-exploratory-data-analysis-ml-ppol670.html#37) -->

<!-- - **Practice**: -->
<!--   - The following provides a [walkthrough](Lectures/week_12/walkthrough/interpretable-ml-walkthrough.html) of the concepts covered in this lecture. The materials can be found  [here](Lectures/week_12/week-12-walkthrough.zip) as a `.zip`. -->

<br>

### _Week 13_: **Applications in Unsupervised Learning**

<!-- ### _Week 11_: [Applications in Unsupervised Learning](Lectures/week_11/lecture-week-11-applications-unsupervised-ml-ppol670.html) -->
<!-- - **Table of Contents**: -->
<!--   + [Unsupervised Learning](http://ericdunford.com/ppol670/Lectures/week_11/lecture-week-11-applications-unsupervised-ml-ppol670.html#3) -->
<!--   + [K Means Clustering](http://ericdunford.com/ppol670/Lectures/week_11/lecture-week-11-applications-unsupervised-ml-ppol670.html#7) -->
<!--   + [Hierarchical Clustering](http://ericdunford.com/ppol670/Lectures/week_11/lecture-week-11-applications-unsupervised-ml-ppol670.html#19) -->

<!-- - **Practice**: -->
<!--   - The following provides a [walkthrough](Lectures/week_11/walkthrough/unsupervised-learning-walkthrough.nb.html) of the concepts covered in this lecture. The materials can be found  [here](Lectures/week_11/week_11-walkthrough.zip) as a `.zip`. -->

<br>



