---
title: "PPOL670"
output: 
  html_document:
    css: style.css
    theme: united
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
---

```{r,include=F}
# Packages for reading in the tables
require(tidyverse)
require(kableExtra)
```



<center><h1>Introduction to Data Science</h1></center>
<center>**Spring 2020**</center>
<center>**Georgetown University**</center>

<br>



# Course Outline {.tabset .tabset-fade .tabset-pills}

## Course Schedule

<br>
<br>


```{r,echo=F,message=F,results='asis'}
read_csv('Syllabus/spring-2020-timeline.csv') %>%
  filter(Topic != "No Class") %>% 
  {.[is.na(.)]="";.} %>% 
  kable(., "html", booktabs = T,align = c('c','l','l')) %>%
  kable_styling(latex_options = c("striped", 
                                  "scale_down","repeat_header"),
                full_width = T,font_size = 24) %>% 
  column_spec(1,width = '.5in') %>% 
  column_spec(2,width = '1in') %>% 
  column_spec(3,width = '4in') 
```

<br>
<br>

---

<br>
<br>


## Syllabus

<br>
<br>

- Access an [**Online Version**](Syllabus/PPOL670-Introduction-to-DS_Syllabus.html)
- [**Download**](Syllabus/PPOL670-Introduction-to-DS_Syllabus.pdf) a `.pdf` version locally. 

<br>
<br>

---

<br>
<br>

## Readings

<br>

List of [**all required and suggested readings**](Readings/reading-list.html)

<br>
<br>

---

<br>
<br>

## Project

<br>

Data science is an applied field and therefore, it is important that you understand how to conduct a complete analysis from collecting data, to cleaning and analyzing it, to presenting your findings. Toward this end, students are required to complete an independent data science project, _applying concepts learned throughout the course_. The project is composed of three parts: a 2 page project proposal, an in-class presentation, and a 12-page project report.

More information regarding the final project will be circulated during class on Week 8. 

<!-- <br> -->

<!-- <center>[**Project Overview**](Assignments/Project/project_overview.pdf)</center> -->

<!-- <center>[**Presentation Assessment Sheet**](Assignments/Project/project_presentation_assessment_sheet.pdf)</center> -->

<!-- <center>[**Final Project Rubric**](Assignments/Project/final_project_rubric.pdf)</center> -->



<br>
<br>

---

<br>
<br>


## Installation

<br>
<br>

The following are installation instructions for `R`, `RStudio`, and `Git`.

<br>
<br>

**`R` Software**

To install `R`, download `R` from CRAN via the following:
 
- Windows: https://cran.r-project.org/bin/windows/base/ (Links to an external site.)
- Mac: https://cran.r-project.org/bin/macosx/
 
To install `RStudio`, download from the following (scroll to the bottom):

- https://www.rstudio.com/products/rstudio/download/ (Links to an external site.)

Video walkthroughs:

- [Video Walkthrough](https://www.dropbox.com/s/7af77pyb1wh224w/R-startup.mp4?dl=0) - [_Put together by former TA (Nathan Lovin)_]
- RStudio Videos
  - [Install `R`](https://player.vimeo.com/video/203516510)
  - [Install `RStudio`](https://player.vimeo.com/video/203516968)

<br>
<br>

**Git** 

- Windows/Mac: https://git-scm.com/downloads (Links to an external site.)
    + Note that git comes pre-installed on the macOS operating system. 

Video walkthroughs:
 
 - [TA Video Walkthrough: Using Git with RStudio](https://www.dropbox.com/s/qdw38hbvy0kj630/git.mp4?dl=0) - [_Put together by former TA (Nathan Lovin)_]


<br>
<br>

---

<br>
<br>


## Contact

<br>

#### [Eric Dunford](http://ericdunford.com/) (Professor)

- **Office**: 404 Old North
- **Office Hours**: Tuesdays 1pm to 3pm
- **Email**: [eric.dunford@georgetown.edu](mailto:eric.dunford@georgetown.edu)

<br>

#### [Trey Billing](https://www.treybilling.com/) (Teaching Assistant)

- **Office Hours**: by appointment -- [book Trey via calendly](https://calendly.com/billingtt/ppol670)
- **Email**: [tb990@georgetown.edu](mailto:tb990@georgetown.edu)

<br>

#### Class

- **Slack**: class communications will primarily take place through [Slack](https://get.slack.help/hc/en-us/articles/218080037-Getting-started-for-new-members). Please follow the [**_invite link_**](https://join.slack.com/t/ppol670introt-vcq1395/shared_invite/enQtODgyMDYyNDQ4ODgyLTFjMTJiZmJlNGQzMTkyNDMwNmUyYjZjZWQxOGIwMmIzZTFmYTAyZWEzOWE5ZDNlNjk4MWZjZTE2OTA3Y2ZhOWQ) to be added to the class Slack channel.

<br>
<br>

---

<br>
<br>




# Lecture Materials

<br>
<br>

### _Week 1_: [Reproducibility and Version Control](Lectures/week_01/lecture-week-01.html)

- **Table of Contents**
  + [Reproducibility](http://ericdunford.com/ppol670/Lectures/week_01/lecture-week-01.html#11)
  + [Version Control](http://ericdunford.com/ppol670/Lectures/week_01/lecture-week-01.html#25)

- **Tutorial**
  + [**Using Git with Rstudio**](Lectures/week_01/tutorial-git-with-rstudio.html)

<br>

### _Week 2_: [Introduction to Programming in `R`](Lectures/week_02/lecture-week-02_intro-to-R.html)

- **Table of Contents**
  + [Interacting with `R`](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#3)
  + [Objects](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#20)
  + [Data Structures](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#37)
  + [Accessing Relational Data](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#45)
  + [Operators](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#70)
  + [Functions](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#84)
  + [Packages](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#90)
  + [Importing/Exporting Data](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#99)
  + [Directories](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#114)

- **Practice**
  + [An interactive practice set covering objects, data structures, and operators](https://dunforde.shinyapps.io/week02-interactive-practice-01/)
  + <a href="Lectures/week_02/Practice/week02_practice_objects-and-operators.R" download> An overview of objects and operators (download `.R`)</a>
  + <a href="Lectures/week_02/Practice/week-02_practice-importing-data.html" target="_blank">Practice importing data</a>

<!-- ### _Week 3_: Reproducibility in Practice -->

<!-- - [**Lecture**](Lectures/week_03/lecture-week03-reproducibility-in-practice.html) -->
<!-- - **Practice**: -->
<!--   + <a href="Lectures/week_03/Practice/working_with_iterators_inclass_practice.R" download> In-class practice with Iterators (download `.R`) </a> -->
<!--   + <a href="Lectures/week_03/Practice/rmarkdown_walkthrough.Rmd" download> R Markdown Walkthrough (download `.Rmd`) </a> -->

<!-- ### _Week 4_: Data Wrangling in R -->

<!-- - [**Lecture**](Lectures/week_04/lecture-week-04_data-wrangling.html) -->
<!-- - **Practice**: -->
<!--   + Practice  <a href="Lectures/week_04/practice/week04-practice-01_Data-Wrangling.Rmd" download> wrangling data (download `.Rmd`) </a> then compare your results to the <a href="Lectures/week_04/practice/week04-practice-01_Data-Wrangling_ANSWERS.Rmd" download> answer key (download `.Rmd`) </a> -->
<!--   + Practice <a href="Lectures/week_04/practice/week04-practice-02_joining-reshaping.Rmd" download> joining and reshaping data (download `.Rmd`) </a> then compare your results to the <a href="Lectures/week_04/practice/week04-practice-02_joining-reshaping_ANSWERS.Rmd" download> answer key (download `.Rmd`) </a> -->


<!-- ### _Week 5_: Data Visualization -->

<!-- - [**Lecture**](Lectures/week_05/lecture-week-05_data-visualization.html) -->
<!-- - **Practice**:  -->
<!--   - Practice materials and data for the visualizations generated in the lecture can be downloaded [here](Lectures/week_05/Practice-Week-05-Visualization.zip) as a `.zip`. -->


<!-- ### _Week 6_: Web Scrapping and APIs -->

<!-- - [**Lecture**](Lectures/week_06/lecture-week-06_webscraping-and-apis.html) -->
<!-- - **Practice**:  -->
<!--   - The following practice has you download a dataset on all bilaterial investment treaties. The practice and answer key can be downloaded [here](Lectures/week_06/week_06_practice.zip) as a `.zip`. -->

<!-- ### _Week 7_: Text as Data -->

<!-- - [**Lecture**](Lectures/week_07/lecture-week-07-text-as-data-ppol670.html) -->
<!-- - **Practice**:  -->
<!--   - The following provides a walkthrough of the concepts covered in this lecture. The materials can be found  [here](Lectures/week_07/week_07_practice.zip) as a `.zip`. -->

<!-- ### _Week 8_: Introduction to Statistical Learning -->

<!-- - [**Lecture**](Lectures/week_08/week-08-lecture-supervised-learning.html) -->

<!-- ### _Week 9_: Applications in Supervised Learning (Regression) -->

<!-- - [**Lecture**](Lectures/week_09/lecture-week-09-applications-supervised-ml-regression-ppol670.html) -->
<!-- - **Practice**: -->
<!--   - The following provides a [walkthrough](Lectures/week_09/walkthrough/supervised-ml-regression-walkthrough.html) of the concepts covered in this lecture. The materials can be found  [here](Lectures/week_09/week_09-practice.zip) as a `.zip`.  -->

<!-- ### _Week 10_: Using Spatial Data-->

<!-- ### _Week 11_: Applications in Supervised Learning (Classification) -->

<!-- - [**Lecture**](Lectures/week_10/lecture-week-10-applications-supervised-ml-classification-ppol670.html) -->
<!-- - **Practice**: -->
<!--   - The following provides a [walkthrough](Lectures/week_10/walkthrough/supervised-ml-classification-walkthrough.html) of the concepts covered in this lecture. The materials can be found  [here](Lectures/week_10/week_10-practice.zip) as a `.zip`.  -->


<!-- ### _Week 12_: Applications in Unsupervised Learning -->

<!-- - [**Lecture**](Lectures/week_11/lecture-week-11-applications-unsupervised-ml-ppol670.html) -->
<!-- - **Practice**: -->
<!--   - The following provides a [walkthrough](Lectures/week_11/week_11-practice/unsupervised-learning-walkthrough.nb.html) of the concepts covered in this lecture. The materials can be found  [here](Lectures/week_11/week_11-practice.zip) as a `.zip`.  -->

<!-- ### _Week 13_: Exploratory Data Analysis -->

<!-- - [**Lecture**](Lectures/week_12/lecture-week-12-exploratory-data-analysis-ml-ppol670.html) -->

<!-- <br>  -->


