---
pagetitle: PPOL670 - Introduction to Data Science
output: 
  html_document:
    css: style.css
    theme: united
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
---

```{r,include=F}
# Packages for reading in the tables
require(tidyverse)
require(kableExtra)

# FUTURE UPDATES
# Add an additional data wrangling day (or extend the current line up) to include:
# - case_when()
# - janitor

# Add gghighlight() to data viz day. 
# Add interactive graphics

# No PDFs on assignments. 

# More in-class coding. Slow down so there can be more of this. 
```

<center><h1>Introduction to Data Science</h1></center>
<center><h2>PPOL670-01</h2></center>
<center>**Fall 2020**</center>
<center>**Georgetown University**</center>

<br>



# Course Outline {.tabset .tabset-fade .tabset-pills}

## Course Schedule

<br>
<br>


```{r,echo=F,message=F,results='asis'}
read_csv('Syllabus/fall-2020-timeline.csv') %>%
  # filter(Topic != "No Class") %>% 
  {.[is.na(.)]="";.} %>% 
  kable(., "html", booktabs = T,align = c('c','l','l')) %>%
  kable_styling(latex_options = c("striped", 
                                  "scale_down","repeat_header"),
                full_width = T,font_size = 20) %>% 
  column_spec(1,width = '.5in') %>% 
  column_spec(2,width = '2in') %>% 
  column_spec(3,width = '4in') 
```

---

<br>
<br>

## Virtual Classroom


<br> 


#### [Link to Virtual Zoom Classroom](https://georgetown.zoom.us/j/91598655229)
  
<br>
  
#### [Virtual Office Hours](https://georgetown.zoom.us/j/96444963432) (Tuesdays 9am - 11am)

  
<br>

Recurrent Zoom link can also be found on Canvas. Please contact the professor/TA through Slack if any of the Zoom links break. 

<br>
<br>
<br>
<hr>
<br>

## Syllabus

<br>
<br>

#### Access an [**online version**](Syllabus/PPOL670-Introduction-to-DS_Syllabus.html)/[**`.pdf` version**](Syllabus/PPOL670-Introduction-to-DS_Syllabus.pdf).

<br>
<br>

---

<br>
<br>

## Readings

<br>

#### List of [**all required and suggested readings**](Readings/reading-list.html)

<br>
<br>

---

<br>
<br>


## Project

<br>

Data science is an applied field and therefore, it is important that you understand how to conduct a complete analysis from collecting data, to cleaning and analyzing it, to presenting your findings. Toward this end, students are required to complete an independent data science project, _applying concepts learned throughout the course_. The project is composed of three parts: a 2 page project proposal, an in-class presentation, and a 12-page project report.

**_More information regarding the final project will be circulated during class on Week 8_**



<!-- <br> -->

<!-- **Overview** -->

<!-- <center>[**Project Overview**](Assignments/Project/project_overview.pdf)</center> -->

<!-- <br> -->

<!-- **Rubrics** -->

<!-- <center>[**Presentation Assessment Sheet**](Assignments/Project/project_presentation_assessment_sheet.pdf)</center> -->

<!-- <center>[**Final Project Rubric**](Assignments/Project/final_project_rubric.pdf)</center> -->

<!-- <br> -->

<!-- **Examples of Successful Projects** -->

<!-- - [Kristina Costa - "Analyzing FEMA Pre-Disaster Mitigation Grants"](Assignments/Project/Costa_Kristina_Example_Final_Report.pdf) -->
<!-- - [Liping Wang - "Female Labor Force Participation"](Assignments/Project/Wang_Liping_Example_Final_Report.pdf) -->



<br>
<br>

---

<br>
<br>


## Installation

<br>
<br>

The following are installation instructions for `R` and `RStudio`.

<br>
<br>

**`R` Software**

To install `R`, download `R` from CRAN via the following:
 
- Windows: https://cran.r-project.org/bin/windows/base/ (Links to an external site.)
- Mac: https://cran.r-project.org/bin/macosx/
 
To install `RStudio`, download from the following (scroll to the bottom):

- https://www.rstudio.com/products/rstudio/download/ (Links to an external site.)

Video walkthroughs:

- [Video Walkthrough](https://www.dropbox.com/s/7af77pyb1wh224w/R-startup.mp4?dl=0) - [_Put together by former TA (Nathan Lovin)_]
- RStudio Videos
  - [Install `R`](https://player.vimeo.com/video/203516510)
  - [Install `RStudio`](https://player.vimeo.com/video/203516968)

<!-- <br> -->
<!-- <br> -->

<!-- **Git**  -->

<!-- - Windows/Mac: https://git-scm.com/downloads (Links to an external site.) -->
<!--     + Note that git comes pre-installed on the macOS operating system.  -->

<!-- Video walkthroughs: -->

<!--  - [TA Video Walkthrough: Using Git with RStudio](https://www.dropbox.com/s/qdw38hbvy0kj630/git.mp4?dl=0) - [_Put together by former TA (Nathan Lovin)_] -->


<br>
<br>

---

<br>
<br>


## Contact

<br>

#### [Eric Dunford](http://ericdunford.com/) (Professor)

- **Office**: Bedroom (formerly 404 Old North)
- **Office Hours**: Tuesdays 9:00am to 11:00am (EST) or by appointment
- **Email**: [eric.dunford@georgetown.edu](mailto:eric.dunford@georgetown.edu)

<br>

#### [Maddie Pickens](http://maddieclairepickens.com/) (Teaching Assistant)

- **Office Hours**: [Book via Calendarly](https://calendly.com/maddiepickens)
- **Email**: [mp1595@georgetown.edu](mailto:mp1595@georgetown.edu)

<br>

#### Class

- **Slack**: class communications will primarily take place through [Slack](https://get.slack.help/hc/en-us/articles/218080037-Getting-started-for-new-members). Please follow the [**_invite link_**](https://join.slack.com/t/ppol670-fall-2020/shared_invite/zt-gnbnemyd-9GSNAaEfXtSUgzsnH3KHKA) to be added to the class Slack channel.

<br>
<br>

---

<br>
<br>


# Lecture Materials
_Asynchronous lecture materials will go live approximately one week prior to the scheduled synchronous meeting date._

<br>
<br>

### _Week 1_: **Work Flow and Reproducibility**

<br>

- [**Lecture Slides**](Lectures/week_01/lecture-week-01.html)

  - **_Table of Contents_**
    + [About this Course](http://ericdunford.com/ppol670/Lectures/week_01/lecture-week-01.html#3)
    + [Reproducibility](http://ericdunford.com/ppol670/Lectures/week_01/lecture-week-01.html#14)
    + [Interacting with `R`](http://ericdunford.com/ppol670/Lectures/week_01/lecture-week-01.html#14)

<br>


### _Week 2_: **Introduction to Programming in `R`**


<br>

- [**Asynchronous Material**](Lectures/week_02/asynchronous_lecture_material/week-02-async-material.html)

- [**Lecture Slides**](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html)

  - **_Table of Contents_**
    - [Objects](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#3)
    - [Data Structures](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#20)
    - [Indexing](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#28)
    - [Operators](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#57)
    - [Functions](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#70)
    - [Packages](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#76)
    - [Importing and Exporting Data](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#85)
    - [But where is my data exactly?](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#100)

- **Practice**
  + In breakout rooms, let's [practice creating an R project, subsetting and creating data structures in R, and importing/exporting data.](Lectures/week_02/Practice/week-02_practice.html)

<br>

### _Week 3_: **Reproducibility in Practice**

<br>

- [**Asynchronous Material**](Lectures/week_03/asynchronous_lecture_material/week-03-async-material.html)


- [**Lecture Slides**](Lectures/week_03/lecture-week03-reproducibility-in-practice.html)

  - **Table of Contents**
    + [Iteration](http://ericdunford.com/ppol670/Lectures/week_03/lecture-week03-reproducibility-in-practice.html#3)
    + [Control Statements](http://ericdunford.com/ppol670/Lectures/week_03/lecture-week03-reproducibility-in-practice.html#22)
    + [`R` Markdown](http://ericdunford.com/ppol670/Lectures/week_03/lecture-week03-reproducibility-in-practice.html#26)
    + [Generating Reproducible Examples](http://ericdunford.com/ppol670/Lectures/week_03/lecture-week03-reproducibility-in-practice.html#50)

- **Practice**:
  + _Breakout 1_: [Piecing together the code](Lectures/week_03/Practice/breakout01_piecing-together-code.html)
  + <a href="Lectures/week_03/Practice/rmarkdown_walkthrough.Rmd.zip" download> R Markdown Walkthrough (download `.Rmd`) </a>
  + _Breakout 2_: [Generate a Reproducible Example](Lectures/week_03/Practice/breakout02_reprex_practice.html)
  

<br>

### _Week 4_: **Data Wrangling in R**


<br>

- [**Asynchronous Material**](Lectures/week_04/asynchronous_lecture_material/week4-async-material.html)


- [**Lecture Slides**](Lectures/week_04/lecture-week-04_data-wrangling.html)
  - **Table of Contents**
    + [Data Wrangling](http://ericdunford.com/ppol670/Lectures/week_04/lecture-week-04_data-wrangling.html#3)
    + [Tidy Data](http://ericdunford.com/ppol670/Lectures/week_04/lecture-week-04_data-wrangling.html#9)
    + [Manipulating Data](http://ericdunford.com/ppol670/Lectures/week_04/lecture-week-04_data-wrangling.html#17)
    + [Piping](http://ericdunford.com/ppol670/Lectures/week_04/lecture-week-04_data-wrangling.html#60)
    + [Joining Data](http://ericdunford.com/ppol670/Lectures/week_04/lecture-week-04_data-wrangling.html#67)
    + [Reshaping Data](http://ericdunford.com/ppol670/Lectures/week_04/lecture-week-04_data-wrangling.html#85)
    
- **Practice**:
  + _Breakout 1_: <a href="Lectures/week_04/walkthrough/breakout01_manipulating-conflict-data.Rmd.zip" download> <strong>Manipulating conflict event data </strong> </a>
  + _Breakout 2_: <a href="Lectures/week_04/walkthrough/breakout02_joining-reshaping.Rmd.zip" download> <strong> Practice joining and reshaping data</strong> </a>

<br>

### _Week 5_: **Data Visualization**

<br>

- [**Asynchronous Material**](Lectures/week_05/asynchronous_lecture_material/week5-async-material.html)

- [**Lecture Slides**](http://ericdunford.com/ppol670/Lectures/week_05/lecture-week-05_data-visualization.html)
  - **Table of Contents**
    + [The Components of Data Visualization](http://ericdunford.com/ppol670/Lectures/week_05/lecture-week-05_data-visualization.html#2)
    + [Grammar of Graphics](http://ericdunford.com/ppol670/Lectures/week_05/lecture-week-05_data-visualization.html#15)
  

- **Practice**:
  - _Walkthrough_: let's visualize the [Gapminder](https://www.gapminder.org/) dataset in a live coding walkthrough. 
    + <a href="Lectures/week_05/synchronous-materials/gapminder_visualization_practice.Rmd.zip" download><strong>Download the notebook</strong></a> (`.rmd` as a `.zip`)
    + [**Answers**](Lectures/week_05/synchronous-materials/gapminder_visualization_practice_ANSWERS.nb.html) to the notebook. 
    
    
- **Third Party Visualization Packages** that build off of `ggplot2`.
    + [`ggthemes`](https://jrnold.github.io/ggthemes/reference/index.html): for great theme and color schemes. 
    + [`scico`](https://github.com/thomasp85/scico): for beautiful continuous color schemes. 
    + [`gghighlight`](https://cran.r-project.org/web/packages/gghighlight/vignettes/gghighlight.html): for highlighting values or subgroups in a ggplot2 plot. 
    + [`ggrepel`](https://cran.r-project.org/web/packages/ggrepel/vignettes/ggrepel.html): for adding text labels to `ggplot2` plots. 
    + [`GGally`](https://ggobi.github.io/ggally/reference/index.html): for generating standard (but more complicated to construct) `ggplot2` plots, such as correlation heatmaps and pairs plots. 

<br>

### _Week 6_: **Web Scrapping**

<br>

- [**Asynchronous Material**](Lectures/week_06/asynchronous_lecture_material/week6-async-material.html)

- [**Lecture Slides**](http://ericdunford.com/ppol670/Lectures/week_06/lecture-week-06_webscraping.html)
  - **Table of Contents**:
    + [Functions](http://ericdunford.com/ppol670/Lectures/week_06/lecture-week-06_webscraping.html#2)
    + [Websites](http://ericdunford.com/ppol670/Lectures/week_06/lecture-week-06_webscraping.html#5)
    + [Scraping Content](http://ericdunford.com/ppol670/Lectures/week_06/lecture-week-06_webscraping.html#14)
    + [Building a Scraper](http://ericdunford.com/ppol670/Lectures/week_06/lecture-week-06_webscraping.html#28)
    + [Legality](http://ericdunford.com/ppol670/Lectures/week_06/lecture-week-06_webscraping.html#35)
    + [Dates](http://ericdunford.com/ppol670/Lectures/week_06/lecture-week-06_webscraping.html#40)
    
- **Discussion**: 
  - [Data Acquisition using `R` packages (wrappers on APIs)](Lectures/week_06/synchronous-materials/data-acquisition-using-r-package-apis.html)

- **Practice**:
  + _Breakout 1_: [Scraping Wikipedia for information on Presidential Debates](Lectures/week_06/synchronous-materials/breakout/breakout01_scraping-wikipedia.html)


<br>

### _Week 7_: **Geospatial Data**

<br>

- [**Asynchronous Material**](Lectures/week_07/asynchronous_lecture_material/week-07-async-materials.html)


- [**Lecture Slides**](http://ericdunford.com/ppol670/Lectures/week_07/asynchronous_lecture_material/slides/week-07-introduction-to-geospatial.html)

- **Practice**:
  + _Breakout_: [**Mapping Conflict in Nigeria**](Lectures/week_07/breakout/breakout_mapping-conflict-in-nigeria.nb.html)
  + [**_Answers to the Breakout_**](Lectures/week_07/breakout/breakout_mapping-conflict-in-nigeria_ANSWERS.nb.html)


- **Discussion**:

  + Retrieve country-level and subnational-level shape files for free from [diva-gis.org](https://www.diva-gis.org/).
  + Explore other popular mapping packages in `R`:
    + [**`ggmap`**](https://github.com/dkahle/ggmap): `ggmap` is an R package that makes it easy to retrieve raster map tiles from popular online mapping services like Google Maps and Stamen Maps and plot them using the ggplot2 framework
    + [**`leaflet`**](https://rstudio.github.io/leaflet/): `leaflet` is one of the most popular open-source JavaScript libraries for interactive maps. It’s used by websites ranging from The New York Times and The Washington Post to GitHub and Flickr, as well as GIS specialists like OpenStreetMap, Mapbox, and CartoDB.
    + [**`tidycensus`**](https://walker-data.com/tidycensus/): `tidycensus` is an R package that allows users to interface with the US Census Bureau’s decennial Census and five-year American Community APIs and return tidyverse-ready data frames, optionally with simple feature geometry included.

<br>

### _Week 8_: **Text as Data**

<br>

- [**Asynchronous Material**](Lectures/week_08/asynchronous-lecture-material/week-08-async-materials.html)

<!-- - [**Lecture Slides**](http://ericdunford.com/ppol670/Lectures/week_08/lecture-week-08-text-as-data-ppol670.html) -->

<!--   - **Table of Contents** -->
<!--     + [Strings](http://ericdunford.com/ppol670/Lectures/week_08/lecture-week-08-text-as-data-ppol670.html#3) -->
<!--     + [Tidy Text](http://ericdunford.com/ppol670/Lectures/week_08/lecture-week-08-text-as-data-ppol670.html#20) -->
<!--     + [Sentiment Analysis](http://ericdunford.com/ppol670/Lectures/week_08/lecture-week-08-text-as-data-ppol670.html#49) -->
<!--     + [Topic Models](http://ericdunford.com/ppol670/Lectures/week_08/lecture-week-08-text-as-data-ppol670.html#64) -->

<!-- - **Practice**: -->
<!--   + _Breakout_: [**Exploring Inaugural Presidential Speeches**](Lectures/week_08/breakout/inaugural-speeches.html) -->
<!--     + Download the <a href="Lectures/week_08/Data/inaug_speeches.csv" download><strong>inaug_speeches.csv</strong></a> dataset used in the breakout.  -->


<br>

### _Week 9_: **Introduction to Statistical Learning**

<!-- ### _Week 8_: [Introduction to Statistical Learning](Lectures/week_08/week-08-lecture-supervised-learning.html) -->
<!-- - **Table of Contents** -->
<!--   + [Statistical Learning](http://ericdunford.com/ppol670/Lectures/week_08/week-08-lecture-supervised-learning.html#3) -->
<!--   + [Supervised Learning](http://ericdunford.com/ppol670/Lectures/week_08/week-08-lecture-supervised-learning.html#13) -->
<!--   + [Cross-Validation](http://ericdunford.com/ppol670/Lectures/week_08/week-08-lecture-supervised-learning.html#32) -->
<!--   + [Caret](http://ericdunford.com/ppol670/Lectures/week_08/week-08-lecture-supervised-learning.html#56) -->
<!--   + [Pre-Processing Data](http://ericdunford.com/ppol670/Lectures/week_08/week-08-lecture-supervised-learning.html#64) -->

<br>

### _Week 10_: **Applications in Supervised Learning (Regression)**

<!-- ### _Week 9_: [Applications in Supervised Learning (Regression)](Lectures/week_09/lecture-week-09-applications-supervised-ml-regression-ppol670.html) -->
<!-- - **Table of Contents** -->
<!--   + [Regression Methods](http://ericdunford.com/ppol670/Lectures/week_09/lecture-week-09-applications-supervised-ml-regression-ppol670.html#3) -->
<!--   + [Regression Trees](http://ericdunford.com/ppol670/Lectures/week_09/lecture-week-09-applications-supervised-ml-regression-ppol670.html#33) -->
<!--   + [Bagging and Random Forest](http://ericdunford.com/ppol670/Lectures/week_09/lecture-week-09-applications-supervised-ml-regression-ppol670.html#48) -->

<!-- - **Practice**: -->
<!--   - The following provides a [walkthrough](Lectures/week_09/walkthrough/supervised-ml-regression-walkthrough.html) of the concepts covered in this lecture. The materials can be found  [here](Lectures/week_09/week_09-practice.zip) as a `.zip`. -->

<br>

### _Week 11_: **Applications in Supervised Learning (Classification)**

<!-- ### _Week 10_: [Applications in Supervised Learning (Classification)](Lectures/week_10/lecture-week-10-applications-supervised-ml-classification-ppol670.html) -->

<!-- - **Table of Contents**: -->
<!--   + [Classification](http://ericdunford.com/ppol670/Lectures/week_10/lecture-week-10-applications-supervised-ml-classification-ppol670.html#3) -->
<!--   + [Logistic Regression](http://ericdunford.com/ppol670/Lectures/week_10/lecture-week-10-applications-supervised-ml-classification-ppol670.html#16) -->
<!--   + [K-Nearest Neighbors](http://ericdunford.com/ppol670/Lectures/week_10/lecture-week-10-applications-supervised-ml-classification-ppol670.html#24) -->
<!--   + [Classification Trees](http://ericdunford.com/ppol670/Lectures/week_10/lecture-week-10-applications-supervised-ml-classification-ppol670.html#37) -->
<!--   + [Support Vector Machines](http://ericdunford.com/ppol670/Lectures/week_10/lecture-week-10-applications-supervised-ml-classification-ppol670.html#42) -->


<!-- - **Practice**: -->
<!--   - The following provides a [walkthrough](Lectures/week_10/walkthrough/supervised-ml-classification-walkthrough.html) of the concepts covered in this lecture. The materials can be found  [here](Lectures/week_10/week-10-walkthrough.zip) as a `.zip`. -->

<br>

### _Week 12_: **Interpretable Machine Learning**

<!-- ### _Week 12_: [Exploratory Data Analysis](Lectures/week_12/lecture-week-12-exploratory-data-analysis-ml-ppol670.html) -->

<!-- - **Table of Contents**: -->
<!--   + [Interpretable Machine Learning](http://ericdunford.com/ppol670/Lectures/week_12/lecture-week-12-exploratory-data-analysis-ml-ppol670.html#3) -->
<!--   + [Data Exploration in the Social Sciences](http://ericdunford.com/ppol670/Lectures/week_12/lecture-week-12-exploratory-data-analysis-ml-ppol670.html#37) -->

<!-- - **Practice**: -->
<!--   - The following provides a [walkthrough](Lectures/week_12/walkthrough/interpretable-ml-walkthrough.html) of the concepts covered in this lecture. The materials can be found  [here](Lectures/week_12/week-12-walkthrough.zip) as a `.zip`. -->

<br>

### _Week 13_: **Applications in Unsupervised Learning**

<!-- ### _Week 11_: [Applications in Unsupervised Learning](Lectures/week_11/lecture-week-11-applications-unsupervised-ml-ppol670.html) -->
<!-- - **Table of Contents**: -->
<!--   + [Unsupervised Learning](http://ericdunford.com/ppol670/Lectures/week_11/lecture-week-11-applications-unsupervised-ml-ppol670.html#3) -->
<!--   + [K Means Clustering](http://ericdunford.com/ppol670/Lectures/week_11/lecture-week-11-applications-unsupervised-ml-ppol670.html#7) -->
<!--   + [Hierarchical Clustering](http://ericdunford.com/ppol670/Lectures/week_11/lecture-week-11-applications-unsupervised-ml-ppol670.html#19) -->

<!-- - **Practice**: -->
<!--   - The following provides a [walkthrough](Lectures/week_11/walkthrough/unsupervised-learning-walkthrough.nb.html) of the concepts covered in this lecture. The materials can be found  [here](Lectures/week_11/week_11-walkthrough.zip) as a `.zip`. -->

<br>



