---
title: "PPOL670 | Introduction to Data Science | Week 8"
subtitle: "Exploring Inaugural Presidential Speeches"
output: 
  html_document:
    df_print: paged
    theme: united
    toc: true
    toc_float: true
    toc_depth: 4
---

```{r include=F, paged.print=FALSE}
knitr::opts_chunk$set(warning = F,error = F,message = F)
require(tidyverse)
require(tidytext)
require(topicmodels)
```

# Data 

```{r}
# Read the data in
inaug_dat <- read_csv("inaug_speeches.csv") 

# Clean it 
dat <- 
  inaug_dat %>% 
  transmute(president = str_to_lower(Name) %>% str_replace_all(.," ","_"),
            address = case_when(
                str_detect(`Inaugural Address`,"First") ~ "first",
                str_detect(`Inaugural Address`,"Second") ~ "second",
                str_detect(`Inaugural Address`,"Third") ~ "third",
                str_detect(`Inaugural Address`,"Fourth") ~ "fourth",
                T ~ "first"),
            date = as.Date(Date,"%A, %B %d, %Y"),
            year = lubridate::year(date),
            length = str_count(text),
            text = text)

# Adjust for one problematic date
dat[dat$president=="bill_clinton" & dat$address=="second",]$date = as.Date("1997-01-20")
dat[dat$president=="bill_clinton" & dat$address=="second",]$year = 1997

head(dat)
```

<br><br>

# Questions

<br><br>

#### (1) Do speeches get longer over time?

<br><br>

#### (2) Who was the most/least verbose?

<br><br>

#### (3) Convert text to tidy format.

- tokenize using words as the fundamental unit. 
- Remove all stopwords
- Remove all digits. 

<br><br>


#### (4) What are the 30 most frequent words used _across all inaugural speeches_? Please present this information as a bar plot.

<br><br>

#### (5) Remove the top 30 most common words from the inaugural speech data. 

Treat these as stop words that are particular to inaugural speeches (i.e. every president uses these words).

<br><br>

#### (6) What are the top five words that are most unique to each president's inaugural speech in their first term? Please present this information as a faceted bar graph. 

<br><br>

#### (7) On Average, which president's inaugural speech is most "positive"? Which is most "negative"? 

<br><br>

#### (8) Of presidents elected into a second term of office, are they more positive in their second inaugural speech vis-a-vis their first on average? Plot this information as a graph of your choosing. 

> Note: Don't consider FDR's third and fourth inaugural speeches. 

<br><br>

#### (9) Run a topic model on the inaugural speeches setting `k` (the number of topics you're looking for to `5`). Try and interpret the output. 

<br><br>
