---
title: "Week 3 | Breakout 2 | Generate a Reproducible Example"
output: 
  html_document:
    css: async-page-style.css
    highlight: breezedark
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<br>

# Instructions

Generate a reproducible example using `reprex` and the `datapasta` packages introduced in the asynchronous lecture. **Only use the first five rows of the data** when generating your example. **Please post the reproducible example as a _direct message_ to Professor Dunford in _Slack_**.

<br>

# Data 
We'll use the `iris` dataset (which is a built in dataset in `R`). See `?iris` for more details.

```{r}
dat <- iris
summary(dat)
```


# Code

Please generate a reproducible example of the following code. 

```{r}
new_dat <- 
  data.frame(
    ave_sepal_length = mean(dat$Sepal.Length),
    ave_sepal_width = mean(dat$Sepal.Width),
    ave_petal_length = mean(dat$Petal.Length),
    ave_petal_width = mean(dat$Petal.Width),
    n_species = length(unique(dat$Species))
  )
new_dat
```

