---
title: "Markdown Basics"
output: html_notebook
---

R Markdown is a useful way to present results. The goal is reproducibility: every line of data and code must be located in the document for it to "knit". If anything is missing then the document won't compile. 

What does it mean to "knit" a document: let's knit this one. Press...

- Mac: `command + shift + k`
- Windows: `control + shift + k`

 When doing so, R is going to ask you to save the file somewhere. 
 
 When we compile the document, we get a new file that just has our code words and results. Look where you saved your code. There should be a `.html` file there now. Click on it. 
 
 
There are three ways in which we can compile a document: a PDF (if you have a latex distribution), html, and a word document. To change how a document is compiled, alter the `output` argument in the preamble with one of the three:

- `word_document`
- `pdf_document`
- `html_document`

If you try all three, you'll find all three files types in the folder where you saved this script. 


------------------------------------

Markdown is an easy to use html language. Let's review some of the **basics**.

# Header
## Header
### Header 
#### Header

**Bold** or _italize_ text. Or **_both_**! ~~Even cross things out~~

[Hyperlink to an Rmarkdown Cheatsheet!](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf)

> We can emphasize text and quotes.
List items:

* Important point!
    + And another...
    + And another...
    
Or enumerate:

Things to do:

1. "Do Research"
2. Stare blankly at the wall and wonder why I came to grad school.
3. Eat chipotle.

We can even embed latex equations! Both as it's own thing. 
$$ \hat{y} = \beta_0 + \beta_1x + \beta_2x^2 + \epsilon $$
And as something that is nested in a sentence $\beta\epsilon\epsilon\rho$. The key difference is between two dollar signs `$$` and one `$`. Either way, always make sure to sandwich your equation!

$$ x_1 

$$ x_1 $$

For help using, Markdown refer to the "Markdown Quick Reference" guide: `Help > Markdown Quick Reference`

### Embedding Code

The real point of R markdown is to embed your R code in your working script. Here we can create a code chunk by pressing:

- mac: `alt + command + i`
- windows: `alt + control + i`


In the code chunk, we can do all things R. 


```{r}
x = rnorm(100,.5,3)
hist(x,breaks=10,col="lightblue",border="white")
```


We can even embed values right in our text. "The value of the coefficient that I care about is `r round(x[1],3)`".


There are a background features built into the R code chucks which allow us to customize. 

```{r,echo=F}
# Echo == F means the code chunk won't render.
y = x + rnorm(100,8,100)
```


```{r,fig.height=3}
plot(y,x,pch=16,cex=3,col="darkgrey")
```


```{r,eval=F}
# Eval == F means the code chunk won't be evaluated
5 + 8 + 9
```


There are many sub-options. One can even use a different programming languages.

```{python}
text = "This is an important bit of text."
print(text.split(' '))
```


## Configuring the YAML

Finally there are a bunch of different ways that the preamble can be set up. Different configurations yield different layouts. One can even create their own layout using `css`. 

Try some of these:

For a Table of Contents with a header depth of 2 (i.e. `##`)
```{}
title: "Markdown Basics"
output:
  html_document:
    toc: true
    toc_depth: 2
```

```{}
output:
  pdf_document:
    toc: true
    number_sections: true
    toc_depth: 4
```

Explore Different Themes
```{}
output:
  html_document:
    theme: spacelab
    highlight: espresso
```

