---
title: "Week02 | Objects, Data Structures, and Operators"
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```


## Objects

### Exercise 1

*Let's create a object called `my_vector` that is a vector containing all intergers from 1 to 100*. Make sure to print your object

```{r vector-practice, exercise=TRUE, exercise.eval=TRUE}


```

```{r vector-practice-hint}
# Recall we can make a vector using c()
x <- c(1,2,3,4)

# Or we can do so with a colon to specify a range
x <- 1:4

# Recall we can print an object using 
print(x)

# or just evaluating the object in the console
x
```

### Exercise 2

```{r vector-quiz}
quiz(
  question("How many data points can you store into a single vector?",
    answer("1"),
    answer("1000"),
    answer("1e100"),
    answer("As many as you like",correct = TRUE)
  )
)
```


### Exercise 3

*I have created a vector called `foods` that contains character values of different types of foods. Please subset the vector so that it only contains `bananas`, `grapes`, and `strawberries`, and assign it to the object `fruits`.*


```{r subsetting-vectors, exercise=TRUE, exercise.eval=TRUE,exercise.lines = 5}
foods <- c('bananas','bread','beef','grapes',
           'cheese','strawberries','milk')
```


```{r subsetting-vectors-hint,exercise.lines = 3}
# Recall we can subset a vector by referencing its index position
x <- c(1,200,3,5,2,10)
x[c(1,3,6)]
```

### Exercise 4

*What type of data object is `volcano`? Use the `class()` or `typeof()` function to find out.*

*Note that `volcano` is loaded into the environment. We can see everything loaded into an environment using the `ls()` function.*

```{r class-example, exercise=TRUE, exercise.eval=TRUE,exercise.lines = 5}
# volcano


```


### Exercise 4

*Convert ("coerce") the object `volcano` into a `data.frame` object.*

```{r coerce,exercise=TRUE, exercise.eval=TRUE,exercise.lines = 5}

```


```{r coerce-hint}
# recall that we can coerce object into different classes using the as.___ functions
x = 1
as.character(x)
as.double(x)
as.matrix(x)
```


## Data Frames

### Exercise 1 

*The data set `mtcars` is loaded into your environment. How many rows and columns does the data frame have?*

```{r,echo=F}
mtcars = mtcars
```


```{r dim,exercise=TRUE, exercise.eval=TRUE,exercise.lines = 5}

```


```{r dim-hint}
# Recall we can find out the dimensions of a data frame using the following functions
dim()
nrow()
ncol()
```


### Exercise 2

*Print off the first top 6 entries of (`head()`) or bottom 6 entries (`tail()`) of `mtcars`.*

```{r head,exercise=TRUE, exercise.eval=TRUE,exercise.lines = 5}

```


### Exercise 3

*Subset `mtcars` to only contain the last two columns and for rows 6 through 20.*

```{r subset1,exercise=TRUE, exercise.eval=TRUE,exercise.lines = 5}

```


```{r subset1-hint}
# Recall we can subset a data.frame by referencing its rows and columns
head(cars)
cars[2,2]
cars[1:5,]
```


### Exercise 4

*Subset the `mtcars` data to only contain retain cars with greater than $24$ miles per gallon (`mpg`).*

```{r subset2,exercise=TRUE, exercise.eval=TRUE,exercise.lines = 5}

```

```{r subset2-hint}
# Recall we use $ to pull out specific variables
cars$speed

# And we can use these vectors of values to make logical statements
cars$speed > 15

# And we can subset a data frame with a logical vector
cars[cars$speed > 15, ]
```
