---
title: "Week 4 - Practice 2 - Joining and Reshaping Data"
subtitle: "ANSWER KEY"
output: html_notebook
---

```{r include = FALSE}
knitr::opts_chunk$set(warning=F,error=F,message = F)
```

## Instructions

The following code reads in three data files (on Github) and assign the data to the respective object names `dogs`, `beertax`, and `traffic`.

```{r}
# Load the tidyverse package
require(tidyverse)

# Read in the data (from online)
beertax <- read_csv('https://raw.githubusercontent.com/edunford/ppol670/gh-pages/Lectures/week_04/practice/Data/states-beer-data.csv')
dogs <- read_csv('https://raw.githubusercontent.com/edunford/ppol670/gh-pages/Lectures/week_04/practice/Data/states-dogs-data.csv')
traffic <- read_csv('https://raw.githubusercontent.com/edunford/ppol670/gh-pages/Lectures/week_04/practice/Data/states-traffic-data.csv')
```

Please use the `_join()` functions from the tidyverse (`dplyr`) to complete each question. In addition, some of the questions require you to reshape the data. For these question, please use the `gather()` and `spread()` functions (`tidyr`). Both functions can be acquired by loading the `tidyverse` package


# Questions


## Question 1

#### (a) Print off the column names in each data set. Are there any ID variables that are mutually occurring across the 3 datasets? Any potential issues?

```{r}
colnames(dogs)
```

```{r}
colnames(beertax)
```

```{r}
colnames(traffic)
```

#### (b) Examine the structure (`str()`) of all three datasets. What are the differences across the datasets?


```{r}
str(dogs)
```

```{r}
str(beertax)
```


```{r}
str(traffic)
```


## Question 2

#### (a) Join `beertax` and `dogs` data so that only the states in `dogs` are retained? Assign to the object named `join1`.

```{r}
join1 = right_join(beertax,dogs,by=c('STATE'='state','YEAR'='year'))
join1
```

#### (b) How many entries for the `beer_tax` variable are missing from the `join1` data? Why are these values missing?

```{r}
join1 %>% 
  summarize(missing = sum(is.na(beer_tax)))
```


#### (c) Join `beertax` and `dogs` data so that only the states in `beertax` are retained? Assign to the object named `join2`.

```{r}
join2 = left_join(beertax,dogs,by=c('STATE'='state','YEAR'='year'))
join2
```

#### (d) How many entries for the `dogs_per_1000` variable are missing from the `join2` data? Why are these values missing?

```{r}
join2 %>% 
  summarize(missing = sum(is.na(dogs_per_1000)))
```


## Question 3

#### (a) join the three datasets so that no values are dropped from either the right or left dataset and assign to the object name `fulldat`.

```{r}
fulldat <-
  full_join(dogs,beertax,by=c('state'='STATE','year'='YEAR')) %>% 
  full_join(.,traffic,by=c('state'='state_name','year'='date_year'))
fulldat
```

#### (b) how many observations are missing from the `dogs_per_1000`, `beer_tax`, and `traffic_per_1000` variables in `fulldat`? Why are some values missing in some of the variable columns?

```{r}
fulldat %>% 
  summarize(missing_dogs = sum(is.na(dogs_per_1000)),
            missing_beer = sum(is.na(beer_tax)),
            missing_traffic = sum(is.na(traffic_per_1000)))
```


## Question 4

#### (a) spread the `fulldat` data into a wide format. Where the `year` makes up the columns and `traffic_per_1000` populate the values for each of the spread out columns. Fill all missing values with a `-99`. Assign the output to the object `fulldat2`.

```{r}
fulldat2<-
  fulldat %>% 
  pivot_wider(names_from = year,
              values_from=traffic_per_1000,
              values_fill=list(traffic_per_1000=-99))

fulldat2
```


#### (b) using `replace_na()` replace all missing values with `-99` in `fulldat2`. Overwrite the `fulldat2` object to retain your changes. (Tip: use the documentation to learn how to use `?replace_na()`)

```{r}
fulldat2 <-
  fulldat2 %>% 
  replace_na(list(dogs_per_1000 = -99, beer_tax = -99))
fulldat2
```
