---
title: "Week 4 - Data Wrangling - Practice"
output:
  html_document:
    df_print: paged
    toc: yes
  html_notebook:
    toc: yes
    toc_float: yes
---

# Practice Instructions

For the following practice, let's use the `flights` dataset from the `nycflights13` package. 

```{r}
# Load the packages
require(tidyverse)
require(nycflights13)
```

For information on the dataset, please consult the documentation.
```{r}
?flights
```

Let's look at the data and store it in an object called `dat`.
```{r}
dat <- flights
glimpse(dat)
```


**Please answer the following questions using the `dplyr` functions we discussed in class. Be sure to practice using the pipe `%>%` when combining functions.** Also, recall the cheat sheet for the `dplyr` function in class. Feel free to reference it as you work through the problems. (See the `Resources/` folder for this material to locate it in the project).

# Question 1

How many observations (i.e. flights) are in the data?

```{r}

```


How many variables are there?

```{r}

```


# Question 2

Subset the data so that we only have the flights that arrive in (`dest`)  and depart (`origin`) from "ATL" ("Atlanta") in January 2013. In the subsetted data frame, only include the following variables:

- `origin`
- `dest`
- `carrier`
- `flight`
- `distance`

Save the subsetted data in an object named `atl_flights`.

```{r}

```

How many flights were there in and out of Atlanta in Jan. 2013?
```{r}

```



How many flights were there in and out of Atlanta in Jan. 2013 _by carrier_? Arrange the data in decending order (that is, those carrier with the most flight to those with the least). Which carrier had the most and least flights?
```{r}

```


# Question 3

On average, which airline has the most delayed departures (`dep_delay`) on average?  Order the carriers from those with the most average delays (in minutes) to those with the least. Which carrier does best on average? Which does worst on average?

> **Tip**: drop missing entries with `drop_na()`


```{r}

```



# Question 4

How many flights are there by month? Which month is the slowest (i.e. which month was there the least number of flights)?

```{r}

```

Which is the slowest flying _day_?

```{r}

```

Which is the slowest flying _day_ by month?

```{r}

```


# Question 5

Using what we know of data manipulation, ask a question of the `flights` data and see if you can find the answer (e.g. How often are there flights to Idaho? In what month are the most longest distance flights? What is the longest distance flight in the data, from where to where? ect.)

```{r}

```

