---
title: "Week 4 - Practice 2 - Joining and Reshaping Data"
output: html_notebook
---


## Instructions

Use data located in the `Data` folder to complete the questions below. Please use the `_join()` functions from the tidyverse (`dplyr`) to complete each question. In addition, some of the questions require you to reshape the data. For these question, please use the `gather()` and `spread()` functions (`tidyr`). Both functions can be acquired by loading the `tidyverse` package

```{r}
require(tidyverse)
```


## Question 1

#### (a) Read in `states-dogs-data.csv`, `states-beer-data.csv`, and `states-traffic-data.csv` from the `Data/` folder. Assign the data to the respective object names `dogs`, `beertax`, and `traffic`.

```{r}

```


#### (b) Print off the column names in each data set. Are there any ID variables that are mutually occurring across the 3 datasets? Any potential issues?

```{r}

```


#### (c) Examine the structure (`str()`) of all three datasets. What are the differences across the datasets?

```{r}

```


## Question 2

#### (a) Join `beertax` and `dogs` data so that only the states in `dogs` are retained? Assign to the object named `join1`.

```{r}

```


#### (b) How many entries for the `beer_tax` variable are missing from the `join1` data? Why are these values missing?

```{r}

```


#### (c) Join `beertax` and `dogs` data so that only the states in `beertax` are retained? Assign to the object named `join2`.

```{r}

```


#### (d) How many entries for the `dogs_per_1000` variable are missing from the `join2` data? Why are these values missing?

```{r}

```



## Question 3

#### (a) join the three datasets so that no values are dropped from either the right or left dataset and assign to the object name `fulldat`.

```{r}

```


#### (b) how many observations are missing from the `dogs_per_1000`, `beer_tax`, and `traffic_per_1000` variables in `fulldat`? Why are some values missing in some of the variable columns?

```{r}

```



## Question 4

#### (a) spread the `fulldat` data into a wide format. Where the `year` makes up the spread columns and `traffic_per_1000` populate the values for the spread columns. Fill all missing values with a `-99`. Assign the output to the object `fulldat2`.

```{r}

```

#### (b) using `replace_na()` replace all missing values with `-99` in `fulldat2`. Overwrite the `fulldat2` object to retain your changes. (Tip: use the documentation to learn how to use `?replace_na()`)

```{r}

```

