---
pagetitle: "PPOL670 | Week 12 - Asynchronous lecture materials"
title:  <a href="http://ericdunford.com/ppol670/">Back to Course Website</a> <br><br><center> Interpretable Machine Learning </center>
subtitle: <center> PPOL 670 | Introduction to Data Science <br><br> Lecture Materials for Week 12 </center><br>
author: <center>Professor Eric Dunford (ed769@georgetown.edu) <br> McCourt School of Public Policy, Georgetown University </center>
output: 
  html_document:
    includes: 
      after_body: async-footer.html
    css: async-page-style.css
    highlight: breezedark
    theme: united
---

<br><hr><br>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Learning Objectives 

<br>

**In the Asynchronous Lecture**

- Talk through why **interpreting machine learning models** matters. 
- Delve into model-agnostic ways of determining **variable importance**.
- Explore marginal effects using **partial dependency** and **individual conditional expectation** plots.

<br>

**In the Synchronous Lecture**

- Explore building **surrogate models** to help interpret blackbox models.
- Apply the model interpretability techniques to explore the health care coverage model we built in Week 11.  

<br>

> If you have any questions while watching the pre-recorded material, be sure to **write them down and to bring them up** during the synchronous portion of the lecture.

<br><hr><br>

# Asynchronous Materials {.tabset .tabset-pills}

<br>

_The following tabs contain pre-recorded lecture materials for class this week. Please review these materials prior to the synchronous lecture._

**_Total time_**: Approx. 1 hour and 6 minutes. 

<br>

## _

<br><hr><br>

## Interpretable Machine Learning

### [**Relevant Slides**](http://ericdunford.com/ppol670/Lectures/week_12/lecture-week-12-interpretable-ml-ppol670.html#3)

<iframe src="https://georgetown.hosted.panopto.com/Panopto/Pages/Embed.aspx?id=38054bb1-54c7-400e-98a2-ac6601369856&autoplay=false&offerviewer=true&showtitle=true&showbrand=false&start=0&interactivity=all" height="506" width="900" style="border: 1px solid #464646;" allowfullscreen allow="autoplay"></iframe>

<br><hr><br>

## Variable Importance

### [**Relevant Slides**](http://ericdunford.com/ppol670/Lectures/week_12/lecture-week-12-interpretable-ml-ppol670.html#9)

<iframe src="https://georgetown.hosted.panopto.com/Panopto/Pages/Embed.aspx?id=7d5aa5c5-5831-4c16-84aa-ac660139281e&autoplay=false&offerviewer=true&showtitle=true&showbrand=false&start=0&interactivity=all" height="506" width="900" style="border: 1px solid #464646;" allowfullscreen allow="autoplay"></iframe>

### Code from the video

```{r,eval=F}
# ENTER CODE
```

<br><hr><br>

## Partial Dependency Plots (PDP)

### [**Relevant Slides**](http://ericdunford.com/ppol670/Lectures/week_12/lecture-week-12-interpretable-ml-ppol670.html#24)

<iframe src="https://georgetown.hosted.panopto.com/Panopto/Pages/Embed.aspx?id=006bb8c6-30f9-4571-8d55-ac660141a07e&autoplay=false&offerviewer=true&showtitle=true&showbrand=false&start=0&interactivity=all" height="506" width="900" style="border: 1px solid #464646;" allowfullscreen allow="autoplay"></iframe>

<br><hr><br>


## Individual Conditional Expectation Plots (ICE)

### [**Relevant Slides**](http://ericdunford.com/ppol670/Lectures/week_12/lecture-week-12-interpretable-ml-ppol670.html#29)

<iframe src="https://georgetown.hosted.panopto.com/Panopto/Pages/Embed.aspx?id=6b40ca1e-b6c0-4d97-97bc-ac66014587ba&autoplay=false&offerviewer=true&showtitle=true&showbrand=false&start=0&interactivity=all" height="506" width="900" style="border: 1px solid #464646;" allowfullscreen allow="autoplay"></iframe>

<br><hr><br>
