---
title:  <a href="http://ericdunford.com/ppol670/">Back to Course Website</a> <br><br><center> Introduction to Programming in R </center>
subtitle: <center> PPOL 670 | Introduction to Data Science <br><br> Lecture Materials for Week 2 </center><br>
author: <center>Professor Eric Dunford (ed769@georgetown.edu) <br> McCourt School of Public Policy, Georgetown University<br><br> </center>
output: 
  html_document:
    includes: 
      after_body: async-footer.html
    css: async-page-style.css
    highlight: breezedark
    theme: united
    toc: true
---

<br><hr><br>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Learning Objectives This Week

<br>

**In the Asynchronous Lecture**

- Learn what an **_object_** is in `R` 

- Cover **_data structures_** in `R` and how to **_access_** data points in a data structure.

- Examine how `R` can be used for calculations by covering its mathematical and logical **_operators_**

- Overview of **_packages_** in `R`.

<br>

**In the Synchronous Lecture**

- Talk through **_importing_** and **_exporting_** data using the `readr`, `haven`, and `readxl` packages.

- Touch on **_working directory_** and `R` **_projects_**

<br>

## Tips

In this class, we'll cover a lot of ground. Keep in mind that we'll **revisit these concepts again and again**, so no worries if everything doesn't sink in on your first pass. 

If you have questions while watching the pre-recorded material, be sure to **write them down and to bring them up** during the synchronous portion of the lecture.

<br><hr><br>

# Asynchronous Materials {.tabset .tabset-pills}

<br>

_The following tabs contain pre-recorded lecture materials for class this week. Please review these materials prior to the synchronous lecture._

**_Total time_**: 1 hour, 28 minutes

<br>

## Objects

### [**Relevant Slides**](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#3)

<iframe src="https://georgetown.hosted.panopto.com/Panopto/Pages/Embed.aspx?id=4444753b-2b68-4ceb-8012-ac1d01199109&autoplay=false&offerviewer=true&showtitle=true&showbrand=false&start=0&interactivity=all" height="506" width="900" style="border: 1px solid #464646;" allowfullscreen allow="autoplay"></iframe>

### Code from the video

```{r,eval=F}
# Object 1
x1 <- 4

# Object 2
x2 <- "This is a cat"

# Addition operator works for some classes, but not all
x1 + 1
x2 + 1

# See the class
class(x1)
class(x2)

# Coerce an object to be another class
as.character(x1)
as.numeric(x2)

# removing objects
rm(x2)
```

<br><hr><br>

## Data Structures

### [**Relevant Slides**](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#20)

<iframe src="https://georgetown.hosted.panopto.com/Panopto/Pages/Embed.aspx?id=673b9d7a-9712-43fd-a7a5-ac1d01216811&autoplay=false&offerviewer=true&showtitle=true&showbrand=false&start=0&interactivity=all" height="506" width="900" style="border: 1px solid #464646;" allowfullscreen allow="autoplay"></iframe>

### Code from the video

```{r,eval=F}
# Scalar data types
x1 <- 4
x2 <- "A"
x3 <- T

# Looking up the class
class(x3)

# Looking up the length of an object
length(x1)

# Make vector
v1 <- c(1,2,3,4)
length(v1)

# Only homogenous data types
v2 <- c("A",1,1.4,T)
v2
```


<br><hr><br>

## Indexing

### [**Relevant Slides**](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#28)

<iframe src="https://georgetown.hosted.panopto.com/Panopto/Pages/Embed.aspx?id=e31b55bb-7193-484f-9993-ac1e01054160&autoplay=false&offerviewer=true&showtitle=true&showbrand=false&start=0&interactivity=all" height="506" width="900" style="border: 1px solid #464646;" allowfullscreen allow="autoplay"></iframe>

### Code from the video

```{r,eval=F}
# Scalar Data Types -------------------------------------------------------

    4           # Integers
    3.4         # Numeric (double/floats)
    TRUE        # Logical (Boolean)
    "A"         # Character
    factor("A") # Factor
 

# Data Structures ---------------------------------------------------------

    # Vector
    vec <- c("a","b","c","d")
    vec
  
    # List
    my_list <- list(A = 1:4, B = 1:10)
    my_list
    
    # Matrix
    my_mat <- matrix(0, ncol=2,nrow = 5)
    my_mat
    
    # Data Frame
    my_dat <- data.frame(A = 1:4, B = c("a","b","c","d"))
    my_dat

    

# Index -------------------------------------------------------------------

    vec <- c("a","b","c","d","e")
    vec    
    
    vec[-c(1,5,3)]
    vec[-3:-5]
    
    
    data$dist
    data[, 2]
    
    
    data[ 45:50 , c("speed","dist")]
    
    head(data,5)
    tail(data,5)
```


<br><hr><br>

## Operators

### [**Relevant Slides**](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#57)

<iframe src="https://georgetown.hosted.panopto.com/Panopto/Pages/Embed.aspx?id=4fa5f030-80ed-4b28-a3a8-ac1d013d11fc&autoplay=false&offerviewer=true&showtitle=true&showbrand=false&start=0&interactivity=all" height="506" width="900" style="border: 1px solid #464646;" allowfullscreen allow="autoplay"></iframe>

### Code from the video

```{r,eval=F}




# Math Operators ----------------------------------------------------------

# + addition

    4 + 2
  
    # The addition operator is really function under the hood
    `+`(4,2) 
    
    # You cant' add all data types.
    "a" + 2
  
# - subtraction
    
    4 - 2

# / division
  
    4/5

# * multiplication
    
    4*5

# ^ Exponentiation
    
    2^3
  
    
# Built in Math Functions ---------------------------------
  
  # There are many statistical and mathematical functions in R. They operate the
  # same as any function and we'll see them in used in different ways in this
  # class.
  
  # log()
  # exp()
  # mean()
  # median()


# Logical Operators -------------------------------------------------------

  x <- 3
  y <- 4
  
  # Logical statements
  x == y      # equals to
  x != y      # does not equal
  x >= y      # greater than or equal to
  x <= y      # less than or equal to
  x > y       # greater than
  x < y       # less than
  
  # Compound logical statements
  
  # And operator: & -- if both left and right side of the statement are true,
  # then return true, else false
  x == 3 & y ==4
  x == 3 & y ==1
  
  
  # Or operator: | -- if one or the other side of the statement is true,
  # then return true, else false if both are false
  x == 3 | y ==1
  

# Vectorization -----------------------------------------------------------

  # Perform operations to all values to a vector simultaneously  
  x <- 1:10
  
  x + 5
  x * 100
  
  
  # Vectorize works for logical operators as well
  x > 5
  x != 5 & x > 3

  
# Subsetting --------------------------------------------------------------

  # We can use vectors of boolean values to subset vectors.
  x[x==5]
  x[x > 5]
  x[x != 5 & x > 3] # Returns a subset
  x[x != 5 | x > 3] # Returns everything!
  
  

# Misc --------------------------------------------------------------------

  
  # Note that we can store vector of logical values to it's own object
  b1 <- x != 5 & x > 3
  # And use it later
  x[b1]
  
  # Also note that we can coerce a logical vector to be numeric, which turns it
  # into an indicator (0,1) variable
  as.numeric(b1)
```

<br><hr><br>

## Functions & Packages

### [**Relevant Slides**](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#70)

<iframe src="https://georgetown.hosted.panopto.com/Panopto/Pages/Embed.aspx?id=1c8fcd23-0521-4496-880b-ac1d0144869d&autoplay=false&offerviewer=true&showtitle=true&showbrand=false&start=0&interactivity=all" height="506" width="900" style="border: 1px solid #464646;" allowfullscreen allow="autoplay"></iframe>

### Code from the video

```{r,eval=F}
# What are functions? ------------------------

    # We've already run into a number of functions
    c() # generate a vector
    as.character() # coerce to character class
    data.frame() # create a data frame
    # etc. ...
    
    
    # Functions are denoted by a word and parentheses
    # "word" with () = word()
    
    # Functions take arguments
    # word(arg1 = 1,arg2 = 2)
    
    # And generate some kind of output
    # output <- word(arg1 = 1,arg2 = 2)
    
    # Functions are objects like any other object in R. 
    
    # Functions are ways to house code. 


# Base Functions  ------------------------
    
    # Base functions come pre-installed on R. When you turn R on, you have
    # access to all base functions,


# Function Documentation  ------------------------

    # We can use `?` to look at a functions documentation
    ?mean()

    
# Why we import packages?  ------------------------

    # We import packages because we get additional functionality that is not
    # native to the R programming environment

# Installing packages  ------------------------
    
    # We can install packages using the `install.packages()` function
    install.packages("ggplot") # We simply need to provide the name of the function. 
    

# Importing Packages ------------------------------------------------------

    # Once a package is installed, we can now import the package and all it's
    # functions using `library()` or `require()`
    library(ggplot2)
    require(ggplot2)
    
    

# Referencing Package Functions without installing the pacakge ------------

    # We can reference package functions without install the package using `::`
    # operator. Specifically, we use <package_name>::<function_name>
    ggplot2::ggplot()
```

<br><hr><br>

#  Practice 

<br>

These exercises are designed to help you reinforce your grasp of the concepts covered in the asynchronous lecture material. 

### [Objects, Data Structures, and Operators](https://dunforde.shinyapps.io/week02-interactive-practice-01/)

<br><hr><br>

# [Feedback](https://forms.gle/E9WQnmYHPaXUpvnW9)

The following survey asks you quick questions regarding the usefulness of the asynchronous lecture materials. Feedback will be used to modify aspect of the asynchronous materials moving forward.

<br><hr><br>

# Synchronous Materials 

## [Lecture Slides](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html)

- **_Table of Contents_**
  - [Objects](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#3)
  - [Data Structures](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#20)
  - [Indexing](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#28)
  - [Operators](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#57)
  - [Functions](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#70)
  - [Packages](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#76)
  - [Importing and Exporting Data](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#85)
  - [But where is my data exactly?](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#100)

<br><hr><br>

<br>


