---
title:  <a href="http://ericdunford.com/ppol670/">Back to Course Website</a> <br><br><center> Introduction to Programming in R </center>
subtitle: <center> PPOL 670 | Introduction to Data Science <br><br> Lecture Materials for Week 2 </center><br>
author: <center>Professor Eric Dunford (ed769@georgetown.edu) <br> McCourt School of Public Policy, Georgetown University </center>
output: 
  html_document:
    includes: 
      after_body: async-footer.html
    css: async-page-style.css
    highlight: breezedark
    theme: united
---

<br><hr><br>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Learning Objectives This Week

<br>

**In the Asynchronous Lecture**

- Learn what an **_object_** is in `R` 

- Cover **_data structures_** in `R` and how to **_access_** data points in a data structure.

- Examine how `R` can be used for calculations by covering its mathematical and logical **_operators_**

- Overview of **_functions_** and **_packages_** in `R`.

<br>

**In the Synchronous Lecture**

- Talk through **_importing_** and **_exporting_** data using the `readr`, `haven`, and `readxl` packages.

- Touch on setting up a **_working directory_** and `R` **_projects_**

<br>

## Tips

For this first day of class, we'll cover a lot of ground. Keep in mind that we'll **revisit these concepts again and again**, so no worries if everything doesn't sink in on your first pass. 

If you have questions while watching the pre-recorded material, be sure to **write them down and to bring them up** during the synchronous portion of the lecture.


<br><hr><br>

# Asynchronous Materials {.tabset .tabset-pills}

<br>

_The following tabs contain pre-recorded lecture materials for class this week. Please review these materials prior to the synchronous lecture._

**_Total time_**: 1 hour, 28 minutes

<br>

## Objects

### [**Relevant Slides**](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#3)

<iframe src="https://georgetown.hosted.panopto.com/Panopto/Pages/Embed.aspx?id=4444753b-2b68-4ceb-8012-ac1d01199109&autoplay=false&offerviewer=true&showtitle=true&showbrand=false&start=0&interactivity=all" height="506" width="900" style="border: 1px solid #464646;" allowfullscreen allow="autoplay"></iframe>

### Code from the video

```{r,eval=F}
# Object 1
x1 <- 4

# Object 2
x2 <- "This is a cat"

# Addition operator works for some classes, but not all
x1 + 1
x2 + 1

# See the class
class(x1)
class(x2)

# Coerce an object to be another class
as.character(x1)
as.numeric(x2)

# removing objects
rm(x2)
```

<br><hr><br>

## Data Structures

### [**Relevant Slides**](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#20)

<iframe src="https://georgetown.hosted.panopto.com/Panopto/Pages/Embed.aspx?id=673b9d7a-9712-43fd-a7a5-ac1d01216811&autoplay=false&offerviewer=true&showtitle=true&showbrand=false&start=0&interactivity=all" height="506" width="900" style="border: 1px solid #464646;" allowfullscreen allow="autoplay"></iframe>

### Code from the video

```{r,eval=F}
# Scalar data types
x1 <- 4
x2 <- "A"
x3 <- T

# Looking up the class
class(x3)

# Looking up the length of an object
length(x1)

# Make vector
v1 <- c(1,2,3,4)
length(v1)

# Only homogenous data types
v2 <- c("A",1,1.4,T)
v2
```


<br><hr><br>

## Indexing

### [**Relevant Slides**]()

[PLUG IN VIDEO HERE + Link to slides]

### Code from the video

```{r,eval=F}
# Scalar Data Types -------------------------------------------------------

    4           # Integers
    3.4         # Numeric (double/floats)
    TRUE        # Logical (Boolean)
    "A"         # Character
    factor("A") # Factor
 

# Data Structures ---------------------------------------------------------

    # Vector
    vec <- c("a","b","c","d")
    vec
  
    # List
    my_list <- list(A = 1:4, B = 1:10)
    my_list
    
    # Matrix
    my_mat <- matrix(0, ncol=2,nrow = 5)
    my_mat
    
    # Data Frame
    my_dat <- data.frame(A = 1:4, B = c("a","b","c","d"))
    my_dat

    

# Index -------------------------------------------------------------------

    vec <- c("a","b","c","d","e")
    vec    
    
    vec[-c(1,5,3)]
    vec[-3:-5]
    
    
    data$dist
    data[, 2]
    
    
    data[ 45:50 , c("speed","dist")]
    
    head(data,5)
    tail(data,5)
```


<br><hr><br>

## Operators

### [**Relevant Slides**](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#53)

<iframe src="https://georgetown.hosted.panopto.com/Panopto/Pages/Embed.aspx?id=4fa5f030-80ed-4b28-a3a8-ac1d013d11fc&autoplay=false&offerviewer=true&showtitle=true&showbrand=false&start=0&interactivity=all" height="506" width="900" style="border: 1px solid #464646;" allowfullscreen allow="autoplay"></iframe>

### Code from the video

```{r,eval=F}
# ENTER CODE
```

<br><hr><br>

## Functions & Packages

### [**Relevant Slides**](http://ericdunford.com/ppol670/Lectures/week_02/lecture-week-02_intro-to-R.html#66)

<iframe src="https://georgetown.hosted.panopto.com/Panopto/Pages/Embed.aspx?id=1c8fcd23-0521-4496-880b-ac1d0144869d&autoplay=false&offerviewer=true&showtitle=true&showbrand=false&start=0&interactivity=all" height="506" width="900" style="border: 1px solid #464646;" allowfullscreen allow="autoplay"></iframe>

### Code from the video

```{r,eval=F}
# ENTER CODE
```

<br><hr><br>

#  Practice 

<br>

These exercises are designed to help you reinforce your grasp of the concepts covered in the asynchronous lecture material. 

### **[Objects, Data Structures, and Operators](https://dunforde.shinyapps.io/week02-interactive-practice-01/)**

<br>


