---
pagetitle: "PPOL670 | Week 3 - Asynchronous lecture materials"
title:  <a href="http://ericdunford.com/ppol670/">Back to Course Website</a> <br><br><center> Reproducibility in Practice </center>
subtitle: <center> PPOL 670 | Introduction to Data Science <br><br> Lecture Materials for Week 3 </center><br>
author: <center>Professor Eric Dunford (ed769@georgetown.edu) <br> McCourt School of Public Policy, Georgetown University </center>
output: 
  html_document:
    includes: 
      after_body: async-footer.html
    css: async-page-style.css
    highlight: breezedark
    theme: united
---

<br><hr><br>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Learning Objectives 

<br>

**In the Asynchronous Lecture**


- Continue learning some `R` programming basics:
  - Iterating over objects using `for` and `while` loops;
  - How to save data manipulations from loops using contains;
  - Conditional code using controls structures.
- Generating reproducible examples using `reprex` and `datapasta`

<br>

**In the Synchronous Lecture**

- Using `RMarkdown` to generate reproducible documents. 

> If you have any questions while watching the pre-recorded material, be sure to **write them down and to bring them up** during the synchronous portion of the lecture.

<br><hr><br>

# Asynchronous Materials {.tabset .tabset-pills}

<br>

_The following tabs contain pre-recorded lecture materials for class this week. Please review these materials prior to the synchronous lecture._

**_Total time_**: XXX

<br>

## `for` Loops

### [**Relevant Slides**]()

[VIDEO HERE]

### Code from the video

```{r,eval=F}
# ENTER CODE
```

<br><hr><br>

## `while` Loops

### [**Relevant Slides**]()

[VIDEO HERE]

### Code from the video

```{r,eval=F}
# ENTER CODE
```

<br><hr><br>

## Containers

### [**Relevant Slides**]()

[VIDEO HERE]

### Code from the video

```{r,eval=F}
# ENTER CODE
```

<br><hr><br>

## Control Structures

### [**Relevant Slides**]()

[VIDEO HERE]

### Code from the video

```{r,eval=F}
# ENTER CODE
```

<br><hr><br>

## Reproducible Examples

### [**Relevant Slides**]()

[VIDEO HERE]

### Code from the video

```{r,eval=F}
# When code breaks --------------------------------------------------------

    # Here is an example of a code chunk with a problem.
    my_data <- cars 
    my_data$new_var <- NA # create a new variable 
    for( i in 1:nrow(my_data)){
      
      # using speed, categorize cars as fast/slow
      if(my_data$speed[1] >= mean(my_data$speed)){
        
        my_data$new_var[i] <- "fast"  
        
      }else{
        
        my_data$new_var[i] <- "slow"
        
      }
      
    }
    
    # Print off
    head(my_data)
    
    # This doesn't look right, all the cars are slow!
    

# Generating a reproducible example of the code ---------------------------------------

    require(reprex) # Load the package
    reprex() # Run the function
    
# Generating a reproducible example of the data ---------------------------------------    
    
    
    require(datapasta) # Load the package

    # Small example
    ex_dat <- head(cars)
    
    # To save output to clipboard
      df_format(ex_dat)
      tribble_format(ex_dat)
    
    # To instantly paste output
      df_paste(ex_dat)
      tribble_paste(ex_dat)
    
# All together now --------------------------------------------------------

    
    # Small representative example of the data
    ex_dat <- head(cars)
    tribble_format(ex_dat)
    
    
    # Tailor the small code example
    
    # Data 
    my_data <- tibble::tribble(
      ~speed, ~dist,
      4,     2,
      4,    10,
      7,     4,
      7,    22,
      8,    16,
      9,    10
    )
    
    
    # Code
    my_data$new_var <- NA # create a new variable 
    for( i in 1:nrow(my_data)){
      
      # using speed, categorize cars as fast/slow
      if(my_data$speed[1] >= mean(my_data$speed)){
        
        my_data$new_var[i] <- "fast"  
        
      }else{
        
        my_data$new_var[i] <- "slow"
        
      }
      
    }
    
    # Print off output
    head(my_data)
    
    
    # Copy the above and then run reprex()
    reprex()
```

<br><hr><br>

# Practice {.tabset .tabset-pills}

<br>

These exercises are designed to help you reinforce your grasp of the concepts covered in the asynchronous lecture material. 

<br>

## _

## Question 1  {.tabset}

<br>

[Enter Question Here]

<br>

### _

### Answer

## Question 2  {.tabset}

<br>

[Enter Question Here]

<br>

### _

### Answer

## Question 3  {.tabset}

<br>

[Enter Question Here]

<br>

### _

### Answer


