---
title: "PPOL 670 | Week 5 | Walthrough"
subtitle: | 
  | Data Visualization Application - Gapminder Data 
output: 
  html_notebook:
    theme: united
    toc: true
    toc_float: true
    toc_depth: 5
---

<br><br>

```{r setup, include= F}
knitr::opts_chunk$set(error=F,warning = F,comment=F)
```



```{r dependencies}
#install.packages("tidyverse")
require(tidyverse) # The tidyverse package covered last time 

# install.packages("ggthemes")
require(ggthemes) # for great visualization colors and themes. 

# install.packages("maps")
require(maps) # for some maps data

# Gapminder data (for example)
#install.packages("gapminder")
require(gapminder)
```

# Data 

The [`Gapminder` dataset](https://www.gapminder.org/) is a famous dataset used by Hans Rosling to visualize development outcomes. The data covers 1952 to 2007 in five year intervals and measures life expectancy, population, and GDP Per Capita. 

```{r}
gapminder %>% head()
```


# Get to know the data through visualization

Each of the following questions are targeted at making sure we understand our data better. A great way to get a "feel" for a dataset is to visualize it. Answer each of the below questions with a (publishable-quality) picture. 

#### 1. How is `lifeExp`, `pop` and `gdpPercap` variables distributed?

Two ways to think about this: one is to plot each figure individually. 
```{r,fig.align="center",fig.width=7,fig.height=4}

```


The better way is to plot all the functions at once using the `pivot_longer()` function from last time with `facet_wrap()`

```{r,fig.align="center",fig.width=10,fig.height=4}

```
We can quickly see that there is large right skews in both `gdpPercap` and `pop`. Let's transform these variables and re-plot.

```{r,fig.align="center",fig.width=10,fig.height=4}

```

Now let's make things look professional!

```{r,fig.align="center",fig.width=10,fig.height=4}

```



#### 2. What's the relationship between economic development and life expectancy? Is the relationship the same for all continents?

```{r,fig.align="center",fig.width=10,fig.height=4}

```

Useful but there are a number of small aesthetic adjustments we could make to really help use distinguish between what is going on. 

```{r,fig.align="center",fig.width=10,fig.height=4}

```

Is the trend the same by continent?

```{r,fig.align="center",fig.width=10,fig.height=4}

```

Generally speaking, it appears so, but it's difficult to hone in on any one continent. There is just a lot going on. Let's consider two alternative ways of presenting this same data. 

Way 1: separate plots using `facet_`

```{r,fig.align="center",fig.width=10,fig.height=4}

```

Way 2: Plot the trends but not the individual data points

```{r,fig.align="center",fig.width=7,fig.height=5}

```

#### 3. Which countries in Africa have the lowest levels of life expectancy?

Two ways we could go about answering a question like this. The first might just be an ordered barplot. Here we might rephrase the question as: "Which countries in Africa have the lowest levels of life expectancy _on average_?"

```{r,fig.align="center",fig.width=7,fig.height=7}

```

Nice, but ordering the factor fields would go a long way. Doing so is easy using the tidy `forcats` package (which is part of the tidyverse), and why we're at it, lets' add a little polish. 

```{r,fig.align="center",fig.width=7,fig.height=7.5}

```

Another way we could approach this is to lay everything out spatially. ggplot with the `maps` package provides a useful way to extract map data on the fly. 

```{r,fig.align="center",fig.width=11,fig.height=7}

```

Our focus is the African continent, so we'll just focus on that portion of the data using the data wrangling principals. 

```{r}

```

Let's plot the map. 

```{r,fig.align="center",fig.width=6,fig.height=7}

```


Now let's fill in the fields on the map using the average life expectancy values.

```{r,fig.align="center",fig.width=6,fig.height=7}

```

What if we wanted to look at how these spatial patterns shifted over time? Not a problem, we just need to tweak the data and plot code a little bit. 

```{r,fig.align="center",fig.width=15,fig.height=15}

```





